---
title: "Untitled"
output: html_document
date: "2024-05-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(dplyr)
library(survival)
library(haven)
library(lubridate)
library(ggsurvfit)
library(gtsummary)
library(readr)
library(consort)
library(splines)
library(readxl)
library(rms)
library(forcats)
library(cmprsk)
library(crrSC)
library(sf)
library(forestplot)
library(gridExtra)
```

## Maps

## Downloading .geoJSON file

Credit to this guide: https://r-graph-gallery.com/325-background-map-from-geojson-format-in-r.html
https://r-graph-gallery.com/327-chloropleth-map-from-geojson-with-ggplot2.html

Map of Chicago

```{r geoJSON}
setwd("~/R/Parker")
my_sf <- read_sf('Boundaries - ZIP Codes.geojson')

```

## Distribution of Scores


```{r ADI}
ADI <- read_csv("2021_ADI.csv")
ADI_crosswalk <-read_excel("11_digit_FIPS_to_ZIP_crosswalk.xlsx")

ADI$FIPS_11_digit <- substr(ADI$FIPS, 1, 11) 
ADI <- left_join(ADI,
                 select(ADI_crosswalk, TRACT, ZIP), 
                 by = c("FIPS_11_digit" = "TRACT") )
ADI <- ADI |> 
  group_by(ZIP) |> 
  mutate(mean_ADI = mean(as.numeric(ADI_NATRANK), na.rm = TRUE))
ADI |> 
  select(ZIP,  ADI_NATRANK, mean_ADI) |> 
  slice_head(n = 5)
ADI <- distinct(ADI, ZIP, .keep_all = TRUE)
rm(ADI_crosswalk)
ADI |> 
  select(ZIP,  ADI_NATRANK, mean_ADI) |> 
  slice_head(n = 5)
```

## SVI


```{r SVI}
SVI <- read_csv("SVI_2022_US.csv")
SVI_crosswalk <- read_excel("11_digit_FIPS_to_ZIP_crosswalk.xlsx")


SVI <- SVI |> subset(STATE == "Illinois")

SVI_crosswalk <- SVI_crosswalk |>  subset(USPS_ZIP_PREF_STATE == "IL")

SVI <- left_join(SVI, select(SVI_crosswalk, TRACT, ZIP), by = c("FIPS" = "TRACT"))
SVI <- SVI |> 
  select(ZIP, RPL_THEMES) 
SVI <- SVI |> mutate(
  RPL_THEMES = case_when(RPL_THEMES < 0 ~ NA, .default = RPL_THEMES),
  SVI_percentile = RPL_THEMES*100)
SVI <- distinct(SVI, ZIP, .keep_all = TRUE)
rm(SVI_crosswalk)
slice_head(SVI, n=10) 


```

## SVM

```{r SVM, include = FALSE}
SVM_ZIP <- read_csv("SVM_ZIPCODE_SCORES_All States_All Counties_2024-01-23.csv")
SVM_ZIP$zipcode <- substr(SVM_ZIP$ZIPCODE, start = 1, stop = 5)

```

## DCI

```{r}
DCI <- read_excel("DCI-2017-2021-Scores-Only.xlsx")
DCI <- DCI |> 
  mutate(zip = as.character(`Zip Code`))
```


## Assign Deprivation Inidicies to ZIP Code


```{r}
my_sf <- left_join(my_sf, 
                              select(SVM_ZIP, ZIPCODE, SVM_SCORE, SVM_PERCENTILE),
                              by = c("zip" = "ZIPCODE"))
my_sf <-left_join(my_sf, SVI, by = c("zip" = "ZIP"))
my_sf <- left_join(my_sf, select(ADI, mean_ADI, ZIP), by = c("zip" = "ZIP"))
my_sf <- left_join(my_sf, select(DCI, `2017-2021 Final Distress Score`, zip), by = c("zip" = "zip"))

my_sf %>%
  ggplot(aes(x = SVI_percentile)) +
  geom_histogram(bins = 10, fill = "skyblue", color = "black")
my_sf %>%
  ggplot(aes(x = SVM_PERCENTILE)) +
  geom_histogram(bins = 10, fill = "purple", color = "black")
my_sf %>%
  ggplot(aes(x = mean_ADI)) +
  geom_histogram(bins = 10, fill = "red", color = "black")
my_sf |> 
  ggplot(aes(x=`2017-2021 Final Distress Score`)) +
  geom_histogram(bins = 10, fill = "green", color = "black")
```

## Producing maps

SVM

```{r SVM map}
SVM_plot <- ggplot(my_sf) +
  geom_sf(aes(fill = SVM_PERCENTILE))+
  theme_void() +
  scale_fill_viridis_c(
    name = "SVM Percentile") +
  theme(legend.title.position = "top",
        legend.title.align = 0.5,
        legend.position = "bottom",
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#f5f5f2", color = NA),
    panel.background = element_rect(fill = "#f5f5f2", color = NA),
    legend.background = element_rect(fill = "#f5f5f2", color = NA),
    plot.title = element_text(
      size = 20, hjust = 0.01, color = "#4e4d47",
      margin = margin(
        b = -0.1, t = 0.4, l = 2,
        unit = "cm"
      )
    ),
    plot.subtitle = element_text(
      size = 15, hjust = 0.01,
      color = "#4e4d47",
      margin = margin(
        b = -0.1, t = 0.43, l = 2,
        unit = "cm"
      )
    )
  )
```

SVI

```{r SVI map}
SVI_plot <- ggplot(my_sf) +
  geom_sf(aes(fill = SVI_percentile))+
  theme_void() +
  scale_fill_viridis_c(
    name = "SVI Percentile") +
  theme(legend.title.position = "top",
        legend.title.align = 0.5,
        legend.position = "bottom",
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#f5f5f2", color = NA),
    panel.background = element_rect(fill = "#f5f5f2", color = NA),
    legend.background = element_rect(fill = "#f5f5f2", color = NA),
    plot.title = element_text(
      size = 20, hjust = 0.01, color = "#4e4d47",
      margin = margin(
        b = -0.1, t = 0.4, l = 2,
        unit = "cm"
      )
    ),
    plot.subtitle = element_text(
      size = 15, hjust = 0.01,
      color = "#4e4d47",
      margin = margin(
        b = -0.1, t = 0.43, l = 2,
        unit = "cm"
      )
    )
  )
```

ADI

```{r ADI map}
ADI_plot <- ggplot(my_sf) +
  geom_sf(aes(fill = mean_ADI))+
  theme_void() +
  scale_fill_viridis_c(
    name = "ADI Percentile") +
  theme(legend.title.position = "top",
        legend.title.align = 0.5,
        legend.position = "bottom",
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#f5f5f2", color = NA),
    panel.background = element_rect(fill = "#f5f5f2", color = NA),
    legend.background = element_rect(fill = "#f5f5f2", color = NA),
    plot.title = element_text(
      size = 20, hjust = 0.01, color = "#4e4d47",
      margin = margin(
        b = -0.1, t = 0.4, l = 2,
        unit = "cm"
      )
    ),
    plot.subtitle = element_text(
      size = 15, hjust = 0.01,
      color = "#4e4d47",
      margin = margin(
        b = -0.1, t = 0.43, l = 2,
        unit = "cm"
      )
    )
  )
```

## DCI

```{r}
DCI_plot <- ggplot(my_sf) +
  geom_sf(aes(fill = `2017-2021 Final Distress Score` ))+
  theme_void() +
  scale_fill_viridis_c(
    name = "DCI Percentile") +
  theme(legend.title.position = "top",
        legend.title.align = 0.5,
        legend.position = "bottom",
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#f5f5f2", color = NA),
    panel.background = element_rect(fill = "#f5f5f2", color = NA),
    legend.background = element_rect(fill = "#f5f5f2", color = NA),
    plot.title = element_text(
      size = 20, hjust = 0.01, color = "#4e4d47",
      margin = margin(
        b = -0.1, t = 0.4, l = 2,
        unit = "cm"
      )
    ),
    plot.subtitle = element_text(
      size = 15, hjust = 0.01,
      color = "#4e4d47",
      margin = margin(
        b = -0.1, t = 0.43, l = 2,
        unit = "cm"
      )
    )
  )
```


## Transplant rate by zip code

```{r transplant by zip}
setwd("~/R/Parker")
cand_thor <- read_sas("cand_thor.sas7bdat", NULL)
cand_zip <- read_sas("canzip2403.sas7bdat", NULL)

cand_thor <- left_join(cand_thor, select(cand_zip,PX_ID, CAN_PERM_ZIP), by = c("PX_ID" = 'PX_ID'))

```

## Subet Cohort

Only include patients added to the waitlist through 2019 or 2022.This allows us to observe everyone for at least 1 year this way.

```{r, echo=FALSE}
cand_thor <- cand_thor |> 
  mutate(eligibility = case_when(
    CAN_AGE_AT_LISTING < 18 ~ "Pediatric candidates",
    CAN_LISTING_DT == CAN_DEATH_DT ~ "Candidate listed and died same day",
  .default = NA))
cand_thor <- cand_thor |> 
  subset(CAN_LISTING_DT > as.Date("2019-01-01") & CAN_LISTING_DT < as.Date("2022-12-31") & WL_ORG != "LU" & is.na(eligibility))

```

## VAD Type characterization

Credit to Kevin L. (https://github.com/kevinlazenby/post_transplant_survival_new_policy/blob/main/function_definitions.Rmd), lines 427 - 460

Durable LVADs (line 867) https://github.com/healthallocate/Heart-Continuous-Score/blob/main/6week_death_data_prep.Rmd


```{r}
cand_thor <- cand_thor |> 
  mutate(VAD = case_when(
        CAN_VAD_TY == 1 ~ paste0("None"),
        CAN_VAD_TY == 2 ~ paste0("LVAD"),
        CAN_VAD_TY == 3 ~ paste0("RVAD"),
        CAN_VAD_TY == 4 ~ paste0("TAH"),
        CAN_VAD_TY == 5 ~ paste0("LVAD+RVAD")),
      
      LVAD = case_when(
        VAD == "LVAD" ~ 1,
        TRUE ~ 0),
      
      durable_LVAD = case_when(
        CAN_VAD1 %in% c('202', '205', '206', '207', '208', '209', '210', 
                        '212', '213', '214', '223', '224', '233', '236', 
                        '239', '240', '312', '313', '314', '315', '316', '319', '322', 
                        '327', '330', '333', '334') | 
          CAN_VAD2 %in%  c('202', '205', '206', '207', '208', '209', '210', 
                        '212', '213', '214', '223', '224', '233', '236', 
                        '239', '240', '312', '313', '314', '315', '316', '319', '322', 
                        '327', '330', '333', '334') ~ 1,
        TRUE ~ 0),

      
      temporary_LVAD = case_when(
        durable_LVAD != 1 & !is.na(CAN_VAD1) ~ 1,
        TRUE ~ 0),
      
      RVAD = case_when(
        VAD == "RVAD" ~ 1,
        (CAN_VAD_TY != 1 & !is.na(CAN_VAD_TY) & 
       !(CAN_VAD1 %in% c('205', '236', '313', '330', '206', '208', '314', 
                         '210', '319', '216', '305', '217', '306', '223', 
                         '312', '224', '316', '230', '324', '231', '325', 
                         '232', '326', '233', '327'))) ~ 1,
        TRUE ~ 0),
      
      TAH = case_when(
        VAD == "TAH" ~ 1,
        TRUE ~ 0),
      
      LVAD_RVAD = case_when(
        VAD == "LVAD+RVAD" ~ 1,
        TRUE ~ 0))
  
      
```


## Intervals

Credit to Kevin Z. in USCRS score (lines 511 - 538)
https://github.com/healthallocate/Heart-Continuous-Score/blob/main/6week_death_data_prep.Rmd


```{r}
 
cand_thor <- cand_thor |> 
  mutate(
    transplant_cand_thor = case_when(!is.na(REC_TX_DT) | CAN_REM_CD == '4' ~ 1,
      TRUE ~ 0),
    start_date = as.Date(CAN_LISTING_DT),
    transplant_date = as.Date(REC_TX_DT),
    removal_date = as.Date(CAN_REM_DT),
    last_active_date = as.Date(CAN_LAST_ACT_STAT_DT),
    last_inactive_date = as.Date(CAN_LAST_INACT_STAT_DT),
    death_date = as.Date(CAN_DEATH_DT),
    death_date_max = pmax(CAN_DEATH_DT, PERS_SSA_DEATH_DT, PERS_OPTN_DEATH_DT, na.rm=T)) %>%
  
  mutate(transplant_time = ifelse(!is.na(transplant_date), 
                                  as.numeric(transplant_date - start_date, units='days'), NA)) %>%
  mutate(death_time = ifelse(!is.na(death_date_max), as.numeric(death_date_max - start_date, units='days'), NA)) %>%
  mutate(removal_time = as.numeric(pmax(removal_date, last_active_date, last_inactive_date, na.rm = T) - 
                                     start_date, units='days')) %>%
  
  
  ## Extra time after removal date in days (can be 0 or negative if no death); cannot be transplanted or dead at removal
  ## i.e. death_date_max - start_date - (removal_date - start_date) = death_date_max - removal_date
  mutate(
    extra_time = case_when(
      transplant_cand_thor != '1' & CAN_REM_CD != '8'  ~ 
        as.numeric(death_time - removal_time, units = 'days'),
      TRUE ~ 0))
```

## Survival time

Survival time defined as time to transplant, death, or removal from waitlist, whichever comes first. Units = days (from previous code)

```{r }
 cand_thor <- 
  cand_thor |> 
  mutate(
    survival_time = pmin(transplant_time, death_time, removal_time, na.rm = TRUE),
    survival_time = case_when(
      survival_time > 1095 ~ 1095,
      .default = survival_time
    ))
```

## Censoring

Death or deterioration on the waitlist coded as the event (1). Transplant is the censoring event (0).

```{r }
cand_thor <- cand_thor |> 
  group_by(PX_ID) |> 
    mutate(status = case_when(
      survival_time == 1095 ~ 0, 
      survival_time == death_time ~ 2,
      CAN_REM_CD == 13 ~ 2,
      CAN_REM_CD == 4  ~ 1,
      CAN_REM_CD == 14 ~ 1,
      CAN_REM_CD == 15 ~ 1,
      CAN_REM_CD == 18 ~ 1,
      CAN_REM_CD == 19 ~ 1,
      CAN_REM_CD == 20 ~ 1,
      CAN_REM_CD == 22 ~ 1,
      CAN_REM_CD == 23 ~ 1,
      .default = 0
      ))

cand_thor$status <- factor(cand_thor$status,
                           levels = c(0,1,2),
                           labels = c("censor",  "transplant","death/deteriorate"))
```

## Transplant and LVAD rates

```{r}
cand_thor$ZIP = substr(cand_thor$CAN_PERM_ZIP, 1,5)
cand_thor <- cand_thor |> 
  group_by(ZIP) |> 
  mutate(transplant_rate = case_when(n() > 0 ~ 100*sum(status == "transplant") / n(),
           .default = 0),
    LVAD_rate = case_when(n() > 0 ~ 100* sum(durable_LVAD == 1) / n(),
      .default = 0
    )
)
my_sf <- left_join(my_sf, select(cand_thor, transplant_rate, LVAD_rate, ZIP), by = c("zip" = "ZIP"))
my_sf <- my_sf |> 
  mutate(transplant_rate = case_when( 
    is.na(transplant_rate) ~ 0, .default = transplant_rate),
    LVAD_rate = case_when(is.na(LVAD_rate) ~ 0, .default = LVAD_rate))
```


## Mapping transplant rate by ZIP
```{r rate map}

transplant_plot <- ggplot(my_sf) +
  geom_sf(aes(fill = transplant_rate))+
  theme_void() +
  scale_fill_viridis_c(
    name = "Transplantation Rate") +
  theme(legend.title.position = "top",
        legend.title.align = 0.5,
        legend.position = "bottom",
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#f5f5f2", color = NA),
    panel.background = element_rect(fill = "#f5f5f2", color = NA),
    legend.background = element_rect(fill = "#f5f5f2", color = NA),
    plot.title = element_text(
      size = 20, hjust = 0.01, color = "#4e4d47",
      margin = margin(
        b = -0.1, t = 0.4, l = 2,
        unit = "cm"
      )
    ),
    plot.subtitle = element_text(
      size = 15, hjust = 0.01,
      color = "#4e4d47",
      margin = margin(
        b = -0.1, t = 0.43, l = 2,
        unit = "cm"
      )
    )
  )

ggplot(my_sf) +
  geom_sf(aes(fill = LVAD_rate))+
  theme_void() +
  scale_fill_viridis_c(
    name = "Rate of Durable LVAD") +
  theme(legend.title.position = "top",
        legend.title.align = 0.5,
        legend.position = "bottom",
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#f5f5f2", color = NA),
    panel.background = element_rect(fill = "#f5f5f2", color = NA),
    legend.background = element_rect(fill = "#f5f5f2", color = NA),
    plot.title = element_text(
      size = 20, hjust = 0.01, color = "#4e4d47",
      margin = margin(
        b = -0.1, t = 0.4, l = 2,
        unit = "cm"
      )
    ),
    plot.subtitle = element_text(
      size = 15, hjust = 0.01,
      color = "#4e4d47",
      margin = margin(
        b = -0.1, t = 0.43, l = 2,
        unit = "cm"
      )
    )
  )
```


## Merging plot

```{r}
grid.arrange(transplant_plot, SVM_plot, SVI_plot, ADI_plot, DCI_plot, ncol = 5)
grid.arrange(SVM_plot, SVI_plot, ADI_plot, DCI_plot, ncol = 4)
```