---
title: "Compare Deprivation Index Heart"
output: html_document
date: "2024-05-23"
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(tidyverse)
library(ggplot2)
library(dplyr)
library(survival)
library(haven)
library(lubridate)
library(ggsurvfit)
library(gtsummary)
library(readr)
library(consort)
library(splines)
library(readxl)
library(rms)
library(forcats)
library(cmprsk)
library(crrSC)
library(sf)
library(forestplot)
library(gridExtra)
library(pROC)
library(coxme)
library(tidycmprsk)
library(xtable)
```

## Loading Data

```{r data}
cand_thor <- read_sas("./pubsaf2403/cand_thor.sas7bdat", NULL)
stathist = read_sas("./pubsaf2403/stathist_thor.sas7bdat")
slice_head(cand_thor,n=5)
```
## ASsigning Scores to Zip codes

## ADI

ADI is coded to the at the block level (12 digit FIPS), whereas the crosswalk document is only coded at the tract level (11 digit FIPS). Need to adjust accordingly. ADI from University of Wisconsin. Crosswalk document from US Department of Housing and Urban Development.

```{r}
ADI <- read_csv("2021_ADI.csv")
ADI_crosswalk <-read_excel("11_digit_FIPS_to_ZIP_crosswalk.xlsx")

ADI$FIPS_11_digit <- substr(ADI$FIPS, 1, 11) 
ADI <- left_join(ADI,
                 select(ADI_crosswalk, TRACT, ZIP), 
                 by = c("FIPS_11_digit" = "TRACT") )
ADI <- ADI |> 
  group_by(ZIP) |> 
  mutate(mean_ADI = mean(as.numeric(ADI_NATRANK), na.rm = TRUE))
ADI |> 
  select(ZIP,  ADI_NATRANK, mean_ADI) |> 
  slice_head(n = 10)
ADI <- distinct(ADI, ZIP, .keep_all = TRUE)
rm(ADI_crosswalk)

```

## SVI

SVI codes missing data as -999. 

```{r}
SVI <- read_csv("SVI_2022_US.csv")
SVI_crosswalk <- read_excel("11_digit_FIPS_to_ZIP_crosswalk.xlsx")

SVI <- left_join(SVI, select(SVI_crosswalk, TRACT, ZIP), by = c("FIPS" = "TRACT"))
SVI <- SVI |> 
  select(ZIP, RPL_THEMES) 
SVI <- SVI |> mutate(
  RPL_THEMES = case_when(RPL_THEMES < 0 ~ NA, .default = RPL_THEMES))
SVI <- SVI |> 
  group_by(ZIP) |> 
  mutate(SVI_percentile = mean(as.numeric(RPL_THEMES*100), na.rm = TRUE),
         median_SVI = median(as.numeric(RPL_THEMES*100), na.rm = TRUE))
SVI <- distinct(SVI, ZIP, .keep_all = TRUE)
rm(SVI_crosswalk)
slice_head(SVI, n=10) 


```

## SVM

```{r}
SVM_ZIP <- read_csv("SVM_ZIPCODE_SCORES_All States_All Counties_2024-01-23.csv")
SVM_ZIP$zipcode <- substr(SVM_ZIP$ZIPCODE, start = 1, stop = 5)
slice_head(SVM_ZIP,n= 5)

```
## DCI

```{r}
DCI <- read_excel("DCI-2017-2021-Scores-Only.xlsx")
DCI <- DCI |> 
  mutate(zip = as.character(`Zip Code`),
         dci_percentile = `2017-2021 Final Distress Score`)
slice_head(DCI, n = 5)
```


## Assign Deprivation Inidicies to Candidate ZIP Code

```{r}
cand_zip <- read_sas("canzip2403.sas7bdat", NULL)
cand_zip$ZIP = substr(cand_zip$CAN_PERM_ZIP,1,5)
cand_zip <- left_join(cand_zip, 
                              select(SVM_ZIP, ZIPCODE, SVM_SCORE, SVM_PERCENTILE),
                              by = c("ZIP" = "ZIPCODE"))
cand_zip <-left_join(cand_zip, SVI, by = c("ZIP" = "ZIP"))
cand_zip <- left_join(cand_zip, select(ADI, mean_ADI, ZIP), by = c("ZIP" = "ZIP"))
cand_zip <- left_join(cand_zip, select(DCI, dci_percentile, zip), by = c("ZIP" = "zip"))
rm(list = c("ADI", "SVI", "SVM_ZIP", "DCI"))

slice_head(cand_zip,n= 5)
```

## Subet Cohort

Only include patients added to the waitlist through 2019 or 2022.This allows us to observe everyone for at least 1 year this way. Only patients that are excluded are patients that are simultaneously listed for heart and lung.

```{r}
cand_pre = cand_thor %>% subset(CAN_LISTING_DT <= as.Date("2018-12-31"))
cand_thor <- cand_thor |> 
  subset(CAN_LISTING_DT >= as.Date("2019-01-01") & CAN_LISTING_DT <= as.Date("2022-12-31") & WL_ORG == "HR")

```

## VAD Type characterization

Credit to Kevin L. (https://github.com/kevinlazenby/post_transplant_survival_new_policy/blob/main/function_definitions.Rmd), lines 427 - 460

Durable LVADs (line 867) https://github.com/healthallocate/Heart-Continuous-Score/blob/main/6week_death_data_prep.Rmd


```{r}
cand_thor <- cand_thor |> 
  mutate(VAD = case_when(
        CAN_VAD_TY == 1 ~ paste0("None"),
        CAN_VAD_TY == 2 ~ paste0("LVAD"),
        CAN_VAD_TY == 3 ~ paste0("RVAD"),
        CAN_VAD_TY == 4 ~ paste0("TAH"),
        CAN_VAD_TY == 5 ~ paste0("LVAD+RVAD")),
      
      LVAD = case_when(
        VAD == "LVAD" ~ 1,
        TRUE ~ 0),
      
      durable_LVAD = case_when(
        CAN_VAD1 %in% c('202', '205', '206', '207', '208', '209', '210', 
                        '212', '213', '214', '223', '224', '233', '236', 
                        '239', '240', '312', '313', '314', '315', '316', '319', '322', 
                        '327', '330', '333', '334') | 
          CAN_VAD2 %in%  c('202', '205', '206', '207', '208', '209', '210', 
                        '212', '213', '214', '223', '224', '233', '236', 
                        '239', '240', '312', '313', '314', '315', '316', '319', '322', 
                        '327', '330', '333', '334') ~ 1,
        TRUE ~ 0),

      
      temporary_LVAD = case_when(
        durable_LVAD != 1 & !is.na(CAN_VAD1) ~ 1,
        TRUE ~ 0),
      
      RVAD = case_when(
        VAD == "RVAD" ~ 1,
        (CAN_VAD_TY != 1 & !is.na(CAN_VAD_TY) & 
       !(CAN_VAD1 %in% c('205', '236', '313', '330', '206', '208', '314', 
                         '210', '319', '216', '305', '217', '306', '223', 
                         '312', '224', '316', '230', '324', '231', '325', 
                         '232', '326', '233', '327'))) ~ 1,
        TRUE ~ 0),
      
      TAH = case_when(
        VAD == "TAH" ~ 1,
        TRUE ~ 0),
      
      LVAD_RVAD = case_when(
        VAD == "LVAD+RVAD" ~ 1,
        TRUE ~ 0))
  
      
```

```{r}
cand_thor <- cand_thor %>% ##cand_thor has same cohort, adds waitlist_end_date column
  mutate(waitlist_end_date = case_when(
    is.na(REC_TX_DT) == FALSE ~ REC_TX_DT, ##look at transplant date first
    is.na(CAN_REM_DT) == FALSE ~ CAN_REM_DT, ##then look at removal date
    is.na(CAN_LAST_INACT_STAT_DT) == FALSE & CAN_LAST_INACT_STAT_DT > CAN_LAST_ACT_STAT_DT ~ CAN_LAST_INACT_STAT_DT,##then inactive date if after last active date
    !is.na(CAN_LAST_ACT_STAT_DT) ~ CAN_LAST_ACT_STAT_DT, ##then last active
    is.na(CAN_LAST_ACT_STAT_DT) & !is.na(CAN_LAST_INACT_STAT_DT) ~ CAN_LAST_INACT_STAT_DT,
    TRUE ~ CAN_LAST_ACT_STAT_DT)) ##if all else fails, use last active date

```

##single registrations - new columns num_list, min_list_date, wait_time, outcome

```{r}
single_registrations <- cand_thor %>%
  group_by(PERS_ID) %>% ##grouping by PERS_ID - ?
  mutate(num_list = n()) %>%
  filter(num_list == 1) %>%
  ungroup() %>%
  mutate(min_list_date = CAN_LISTING_DT,  
         wait_time = waitlist_end_date - min_list_date, ##using end dates established in last code chunk
         outcome = case_when(
           DON_TY == "C" ~ "transplanted",
           !is.na(CAN_REM_CD) ~ "removed/died",  ##do we want all other CAN_REM_CD codes here? or just 8 and 13 (like used to define the outcome below)
           TRUE ~ "censored"
         ))

```

##multiple registrations - new columns num_list, list_type (concurrent or sequential), num_tx; fills missing REC_TX_DT

```{r}
multiple_registrations <- cand_thor %>%
  filter(!PX_ID %in% single_registrations$PX_ID) %>%
  group_by(PERS_ID) %>%
  arrange(PERS_ID, CAN_LISTING_DT) %>%
  mutate(num_list = n())

candidates_w_multiple_registrations <- multiple_registrations %>% pull(PERS_ID) %>% unique() %>% length() ##number of candidates with multiple registrations

avg_num_registrations_multiple <- mean(multiple_registrations %>% group_by(PERS_ID) %>% filter(row_number() == 1) %>% pull(num_list)) ##for candidates with multiple registrations, the average number of registrations

multiple_registrations <- multiple_registrations %>%
  mutate(list_type = case_when(
    CAN_LISTING_DT < lag(waitlist_end_date) ~ "concurrent",
    waitlist_end_date > lead(CAN_LISTING_DT) ~ "concurrent",
    TRUE ~ "sequential")) %>%
  mutate(REC_TX_DT = as.Date(REC_TX_DT)) %>%
  mutate(num_tx = length(unique(na.omit(REC_TX_DT)))) %>%
  fill(REC_TX_DT, .direction='downup') ##this line is causing issues with negative dial_time values
```

```{r}
multiple_registrations <- multiple_registrations[order(multiple_registrations$PERS_ID, multiple_registrations$waitlist_end_date), ] ##orders by PERS_ID then waitlist_end_date

##multiple registrations - number sequential transplants, correct concurrent listing transplant dates

multiple_registrations$transplant_num <- 1 ##new column transplant_num, setting all at 1 for now
for (i in 2:nrow(multiple_registrations)) {
  if (!is.na(multiple_registrations$PERS_ID[i-1]) && 
      !is.na(multiple_registrations$PERS_ID[i]) && 
      !is.na(multiple_registrations$REC_TX_DT[i-1]) && 
      !is.na(multiple_registrations$REC_TX_DT[i]) &&
      multiple_registrations$PERS_ID[i-1] == multiple_registrations$PERS_ID[i] &&
      multiple_registrations$REC_TX_DT[i-1] != multiple_registrations$REC_TX_DT[i]) {
    
    multiple_registrations$transplant_num[i] = multiple_registrations$transplant_num[i-1] + 1
  }
}

for (i in 2:nrow(multiple_registrations)) {
  if (!is.na(multiple_registrations$PERS_ID[i-1]) && 
      !is.na(multiple_registrations$PERS_ID[i]) && 
      multiple_registrations$PERS_ID[i-1] == multiple_registrations$PERS_ID[i] &&
      !is.na(multiple_registrations$transplant_num[i-1]) && 
      multiple_registrations$transplant_num[i-1] != multiple_registrations$transplant_num[i] &&
      multiple_registrations$transplant_num[i-1] != 1) {
    
    multiple_registrations$transplant_num[i] = multiple_registrations$transplant_num[i-1]
  }
}

multiple_registrations$transplant_num[multiple_registrations$list_type == 'sequential'] <- 0 ##when list_type is sequential, transplant_num is set to 0. I still don't understand this step either.

for(i in 1:(nrow(multiple_registrations)-1)) { ##loop from second to lat row
  if(multiple_registrations$PERS_ID[i] == multiple_registrations$PERS_ID[i+1] &
     multiple_registrations$list_type[i] == 'concurrent' & multiple_registrations$list_type[i+1] == 'concurrent' &
     !is.na(multiple_registrations$REC_TX_DT[i]) & !is.na(multiple_registrations$REC_TX_DT[i+1]) &
     multiple_registrations$REC_TX_DT[i] < multiple_registrations$REC_TX_DT[i+1] ) {
    
    multiple_registrations$REC_TX_DT[i] <- multiple_registrations$REC_TX_DT[i+1] ##if PERS_ID is the same as the NEXT row, both have concurrent listings, transplant dates for both are not NA, and transplant date is earlier than the next row's transplant date - update to next row's transplant date
  }}

sequential_lists <- multiple_registrations %>%
  filter(list_type == "sequential") %>%
  mutate(min_list_date = CAN_LISTING_DT,
         wait_time = waitlist_end_date - min_list_date,
         outcome = case_when(
           DON_TY == "C" ~ "transplanted",
           is.na(CAN_REM_CD) == FALSE ~ "removed/died", ##again, should this be all CAN_REM_CD? or just 8 and 13 (since we included the rest in censored)
           TRUE ~ "censored"
         ))  ##same thing as done above for single registrations

max_retransplants <- max(multiple_registrations$transplant_num) ##highest number of transplants for a single person

multiple_registrations <- multiple_registrations %>% ##for candidates with multiple registrations, adds columns min_list_date, wait_time
  group_by(PERS_ID, transplant_num) %>%
  mutate(min_list_date = min(CAN_LISTING_DT, na.rm=T),
         wait_time = waitlist_end_date - min_list_date)

multiple_registrations <- multiple_registrations %>% mutate

collapsed_concurrent_registrations <- NULL ##establishes object to store coming data
for(i in 1:max_retransplants) { ##loop from first row to max_transplants
  
  collapsed_concurrent_registrations <- rbind(collapsed_concurrent_registrations, 
        
  multiple_registrations %>%
    filter(list_type == "concurrent" & transplant_num == i) %>% ##filters for concurrent listings where translpant_num is equal to current value of i (?)
    mutate(DON_TY = ifelse(DON_TY == "", NA, DON_TY),
           last_wait_date = max(waitlist_end_date, na.rm = TRUE)) %>% ##if DON_TY is empty, fill with NA, set last_wait_date to waitlist_end_date (why?)
    fill(REC_TX_DT, .direction = "up") %>%
    fill(DON_TY, .direction = "up") %>%
    fill(DONOR_ID, .direction = "up") %>%
    fill(CAN_REM_CD, .direction = "up") %>% ##fills missing values upwards
    mutate(wait_time = case_when(
      is.na(REC_TX_DT) == FALSE & transplant_num != '0' ~ REC_TX_DT- min_list_date, ### Ignore non-transplanted rows
      TRUE ~ last_wait_date - min_list_date),
      outcome = case_when(
        DON_TY == "C" ~ "DDKT",
        DON_TY == "L" ~ "LDKT",
        is.na(CAN_REM_CD) == FALSE ~ "removed/died", ##same question - CAN_REM_CD should be all? or just 8, 13?
        TRUE ~ "censored")   ##does the same we already did for single registrations and sequential listings
    ) %>%
    select(-c(waitlist_end_date, CAN_LISTING_DT, CAN_REM_DT)) %>% ##drops columns waitlist_end_date, CAN_LISTING_DT, CAN_REM_DT (just for cleanliness?)
    filter(row_number() ==1) %>% ##for concurrent listings, keeps only first row 
    
    mutate(last_wait_date = case_when(
      REC_TX_DT < last_wait_date ~ REC_TX_DT,
      TRUE ~last_wait_date)))}
```

##recreate CAN_LISTING_DT - to fix issue of NA values later on

```{r}
collapsed_concurrent_registrations <- collapsed_concurrent_registrations %>% mutate(CAN_LISTING_DT = min_list_date)
```

##recombine separated data frames

```{r}
cand_thor <- bind_rows(single_registrations %>% ungroup(), 
                          sequential_lists %>% ungroup(), 
                          collapsed_concurrent_registrations %>% ungroup()) ##puts data all back together

cand_thor %>% filter(is.na(CAN_LISTING_DT))

```

## Intervals

Credit to Kevin Z. in USCRS score (lines 511 - 538)
https://github.com/healthallocate/Heart-Continuous-Score/blob/main/6week_death_data_prep.Rmd


```{r intervals}
 
cand_thor <- cand_thor |> 
  mutate(
    transplant_cand_thor = case_when(!is.na(REC_TX_DT) | CAN_REM_CD == '4' ~ 1,
      TRUE ~ 0),
    start_date = as.Date(CAN_LISTING_DT),
    transplant_date = as.Date(REC_TX_DT),
    removal_date = as.Date(CAN_REM_DT),
    last_active_date = as.Date(CAN_LAST_ACT_STAT_DT),
    last_inactive_date = as.Date(CAN_LAST_INACT_STAT_DT),
    death_date = as.Date(CAN_DEATH_DT),
    death_date_max = pmax(CAN_DEATH_DT, PERS_SSA_DEATH_DT, PERS_OPTN_DEATH_DT, na.rm=T)) %>%
  
  mutate(transplant_time = ifelse(!is.na(transplant_date), 
                                  as.numeric(transplant_date - start_date, units='days'), NA)) %>%
  mutate(death_time = ifelse(!is.na(death_date_max), as.numeric(death_date_max - start_date, units='days'), NA)) %>%
  mutate(removal_time = as.numeric(pmax(removal_date, last_active_date, last_inactive_date, na.rm = T) - 
                                     start_date, units='days'))
```

## Survival time

Survival time defined as time to transplant, death, or removal from waitlist, whichever comes first. Units = days (from previous code)

```{r survival time}
 cand_thor <- 
  cand_thor |> 
  mutate(
    survival_time = pmin(transplant_time, death_time, removal_time, na.rm = TRUE),
    survival_time = case_when(
      survival_time > 1095 ~ 1095,
      .default = survival_time
    ))
```

## Censoring

```{r censoring}
cand_thor <- cand_thor |> 
  group_by(PX_ID) |> 
    mutate(status = case_when(
      survival_time == 1095 ~ 0, 
      survival_time == death_time ~ 2,
      survival_time == transplant_time  ~ 1,
      !is.na(CAN_REM_CD) & !(CAN_REM_CD %in% c(4,8)) ~ 3,
      .default = 0
      ))

cand_thor$status <- factor(cand_thor$status,
                           levels = c(0,1,2,3),
                           labels = c("censor",  "transplant","death", "removal"))

cand_thor %>% filter(status != "death/deteriorate" & CAN_REM_CD == 8) %>% select(survival_time, death_time, status)

```

Cause Specific Hazard setup

```{r}
cand_thor <- cand_thor |> 
  mutate(transplant_outcome = case_when(
    status == "censor" ~ 0,
    status == "transplant" ~ 1,
    .default = 0),
    death_outcome = case_when(
      status == "censor" ~ 0,
      status == "death" ~ 1,
      .default = 0
    ))
```

## Figure 1: STROBE/CONSORT Diagram

Inclusion and exclusion criteria Strobe diagram

```{r}
pre_period_pers = cand_pre$PERS_ID

cand_thor <- cand_thor %>%
  group_by(PERS_ID) %>%
  mutate(
    eligibility = case_when(
      PERS_ID %in% pre_period_pers | row_number() != 1 ~ "Not the first record for candidate",
      CAN_AGE_AT_LISTING < 18 ~ "Pediatric candidates",
      CAN_LISTING_DT == CAN_DEATH_DT ~ "Candidate listed and died same day",
      .default = NA
    )
  ) %>%
  ungroup() %>%
  mutate(
    durable_LVAD_table1 = case_when(
      is.na(eligibility) & durable_LVAD == 1 ~ "Durable LVAD",
      is.na(eligibility) & durable_LVAD == 0 ~ "No Durable LVAD",
      .default = NA
    ),
    status_table1 = case_when(
      status == "censor" ~ "On waitlist at end of follow-up",
      status == "transplant" ~ "Received transplant",
      status == "death" ~ "Died",
      status == "removal" ~ "Removed"
    )
  )

txt <- gen_text(rep("Heart transplant candidate registrations, 2019-2022", 19392))
p_cons <- add_box(txt = txt)

# Get unique values from eligibility, excluding "Multiple listings" 
# to avoid additional count for this label
unique_eligibility <- cand_thor$eligibility
unique_eligibility <- c(unique_eligibility, rep("Multiple Listings", 220))
# Combine custom text with the remaining eligibility items


# Generate text without automatic count for "Multiple listings"
txt <- gen_text(unique_eligibility, label = "Excluded", bullet = TRUE)

# Add the customized text to the side box
p_cons <- add_side_box(p_cons, txt = txt)

p_cons <- add_box(p_cons, txt = gen_text(cand_thor$durable_LVAD_table1, label = "Adult candidate registrations"))

## Creating Analytic Sample 
cand_thor <- cand_thor |> 
 subset(is.na(eligibility))

txt <- gen_text(cand_thor$status_table1, bullet = TRUE)
p_cons <- add_box(p_cons, txt = txt, just = "left")


plot(p_cons, grViz = TRUE)

```


## Assigning Deprivation Indices based on Patient's permanent zip


```{r center scores}
heart_SVM <- left_join(cand_thor,
                    select(cand_zip,SVM_SCORE, SVM_PERCENTILE, PX_ID, SVI_percentile, mean_ADI, dci_percentile), 
                    by = c("PX_ID" = "PX_ID"))
```

## Imputing median score



```{r}
heart_SVM <- heart_SVM %>%
  mutate(SVM_PERCENTILE = as.numeric(as.character(SVM_PERCENTILE)), # SVM
         SVI_percentile = as.numeric(as.character(SVI_percentile)), # SVI
         mean_ADI = as.numeric(as.character(mean_ADI)), # ADI
         dci_percentile = as.numeric(as.character(dci_percentile))) # DCI

# Use median score from transplant center listed at for candidates missing score percentile
heart_SVM <- heart_SVM %>%
  group_by(CAN_LISTING_CTR_CD) %>% # Gather by transplant center
  mutate(median_SVM = median(SVM_PERCENTILE, na.rm = T), # Median SVM of each transplant center
         median_SVI = median(SVI_percentile, na.rm = T), # Median SVI of each transplant center
         median_ADI = median(mean_ADI, na.rm = T), # Median ADI of each transplant center
         median_DCI = median(dci_percentile, na.rm = T), # Median DCI of each transplant center
         SVM_PERCENTILE = ifelse(is.na(SVM_PERCENTILE), median_SVM, SVM_PERCENTILE), #conditional to input corresponding composite score value
         SVI_percentile = ifelse(is.na(SVI_percentile), median_SVI, SVI_percentile),  #conditional to input corresponding composite score value
         mean_ADI = ifelse(is.na(mean_ADI), median_ADI, mean_ADI),  #conditional to input corresponding composite score value
         dci_percentile = ifelse(is.na(dci_percentile), median_DCI, dci_percentile))  #conditional to input corresponding composite score value
# Split Percentiles into Deciles
heart_SVM <- heart_SVM %>% 
  group_by(PX_ID) %>%
  mutate(SVM_decile = 
           cut(SVM_PERCENTILE, c(-Inf, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100), labels = c("1st Decile", "2nd Decile", "3rd Decile", "4th Decile", "5th Decile", "6th Decile", "7th Decile", "8th Decile", "9th Decile", "10th Decile")),
           SVI_decile = cut(SVI_percentile, c(-Inf, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100), labels = c("1st Decile", "2nd Decile", "3rd Decile", "4th Decile", "5th Decile", "6th Decile", "7th Decile", "8th Decile", "9th Decile", "10th Decile")),
           ADI_decile = cut(mean_ADI,c(-Inf, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100), labels = c("1st Decile", "2nd Decile", "3rd Decile", "4th Decile", "5th Decile", "6th Decile", "7th Decile", "8th Decile", "9th Decile", "10th Decile")),
         DCI_decile = cut(dci_percentile,c(-Inf, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100), labels = c("1st Decile", "2nd Decile", "3rd Decile", "4th Decile", "5th Decile", "6th Decile", "7th Decile", "8th Decile", "9th Decile", "10th Decile")))
heart_SVM <- heart_SVM |> 
  mutate(svm_decile = as.numeric(SVM_decile),
        svi_decile = as.numeric(SVI_decile),
        adi_decile = as.numeric(ADI_decile),
        dci_decile = as.numeric(DCI_decile))
```




## Covariates

```{r covariates}
heart_SVM <- heart_SVM |> 
  mutate(can_race = case_when(
    CAN_RACE == "8" ~ "White",
    CAN_RACE == "16" ~ "Black or African American",
    CAN_RACE == "32" ~ "American Indian or Alaska Native",
    CAN_RACE == "64" ~ "Asian",
    CAN_RACE == "128" ~ "Native Hawaiian or Other Pacific Islander",
    CAN_RACE == "256" ~ "Arab or Middle Eastern",
    .default = "Other"),
    can_race = as.factor(can_race),
    ethnicity = case_when(
      CAN_ETHNICITY_SRTR == "NLATIN" ~ "Not Latinx",
      CAN_ETHNICITY_SRTR == "LATINO" ~ "Latinx"
      
    ),
    ethnicity = as.factor(ethnicity))
heart_SVM$can_race <- relevel(heart_SVM$can_race, ref = "White")
heart_SVM$ethnicity <- relevel(heart_SVM$ethnicity, ref = "Not Latinx")
heart_SVM <- heart_SVM |> 
  mutate(initial_status = case_when(
           CAN_INIT_STAT %in% c('1010', '1020', '1090', '1110', '2010', '2020', '2090', '2110', '3010', '6010', '6011', '6012', '9010') ~ "Status 1",
           CAN_INIT_STAT %in% c('1030', '1120', '2030', '2120', '6002', '6020', '6030', '9020') ~ "Status 2",
           CAN_INIT_STAT %in% c('1130', '2130', '6040') ~ "Status 3",
           CAN_INIT_STAT %in% c('1140', '2140', '6004') ~ "Status 4",
           CAN_INIT_STAT %in% c('1150', '2150') ~ "Status 5",
           CAN_INIT_STAT %in% c('1160', '2160') ~ "Status 6",
           CAN_INIT_STAT == 2999 ~ "inactive",
           .default = NA
         ),
initial_status = as.factor(initial_status))
```



```{r}
#load other transplant sets for multi listing check
setwd("C:/Users/mhwhi/Downloads")
cand_kipa = read_sas("./pubsaf2403/cand_kipa.sas7bdat")
cand_liin = read_sas("./pubsaf2403/cand_liin.sas7bdat")
```


```{r cars}


#function to check overlapping listings
check_overlap <- function(heart_df, other_df) {
  # Filter for common PERS_ID
  common_ids <- intersect(heart_df$PERS_ID, other_df$PERS_ID)
  
  # Subset datasets to only include common PERS_ID
  filtered_heart_df <- heart_df %>% filter(PERS_ID %in% common_ids)
  filtered_other_df <- other_df %>% filter(PERS_ID %in% common_ids)
  
  # Join heart_df with the filtered other_df for date comparison
  result <- filtered_heart_df %>%
    left_join(filtered_other_df, by = "PERS_ID", suffix = c("_heart", "_other")) %>%
    filter(
      # Case 1: Overlap with removal date
      (CAN_LISTING_DT_heart >= CAN_LISTING_DT_other & 
       CAN_LISTING_DT_heart <= CAN_REM_DT_other) |
      # Case 2: No removal date, listing dates must match or overlap
      (is.na(CAN_REM_DT_other) & CAN_LISTING_DT_heart >= CAN_LISTING_DT_other)
    )
  
  return(result)
}

# Check overlap for both CAND_KIPA and CAND_LIIN
overlap_kipa <- check_overlap(heart_SVM, cand_kipa)
overlap_liin <- check_overlap(heart_SVM, cand_liin)
overlap_kipa <- overlap_kipa %>% select(PERS_ID, WL_ORG_heart, WL_ORG_other)
overlap_kipa = as.data.frame(overlap_kipa %>% group_by(PERS_ID) %>% distinct())
overlap_liin = overlap_liin %>% select(PERS_ID, WL_ORG_heart, WL_ORG_other)
overlap_liin = as.data.frame(overlap_liin %>% group_by(PERS_ID) %>% distinct())
# Combine results
overlap_results <- bind_rows(
  overlap_kipa %>% mutate(Source = "CAND_KIPA"),
  overlap_liin %>% mutate(Source = "CAND_LIIN")
)

#one entry for every patient
consolidated_overlap <- overlap_results %>%
  group_by(PERS_ID) %>%
  dplyr::summarize(WL_ORG_combined = paste(unique(WL_ORG_other), collapse = ", "), .groups = "drop")
consolidated_overlap

heart_SVM <- heart_SVM %>%
  left_join(consolidated_overlap, by = "PERS_ID") %>%
  mutate(
    WL_ORG_other_combined = ifelse(is.na(WL_ORG_combined), "None", WL_ORG_combined)
  )

#indicator variable for multiple listings
heart_SVM = heart_SVM %>% mutate(
  multi_listed = ifelse(!is.na(WL_ORG_combined), 1, 0)
)
```



## Table 1: Cohort Characteristics

```{r table 1}
heart_SVM = heart_SVM %>% mutate(
  Male = ifelse(CAN_GENDER == 'M', 1, 0),
  insurance = case_when(
    CAN_PRIMARY_PAY %in% c(1) ~ "Private insurance",
    CAN_PRIMARY_PAY %in% c(2, 3, 4, 5, 6, 7) ~ "Public insurance",
    CAN_PRIMARY_PAY %in% c(8) ~ "Self",
    CAN_PRIMARY_PAY %in% c(9) ~ "Donation",
    CAN_PRIMARY_PAY %in% c(10) ~ "Free Care",
    CAN_PRIMARY_PAY %in% c(11) ~ "Pending",
    CAN_PRIMARY_PAY %in% c(12) ~ "Foreign Government",
    CAN_PRIMARY_PAY %in% c(13) ~ "Medicare Unspecified",
    CAN_PRIMARY_PAY %in% c(14) ~ "US/State Govt Agency",
    CAN_PRIMARY_PAY %in% c(15) ~ "Unknown",
    TRUE ~ "Other"),
  insurance = as.factor(insurance),
  education = case_when(
    CAN_EDUCATION == 1 ~ "None",
    CAN_EDUCATION == 2 ~ "Grade School (0-8)",
    CAN_EDUCATION == 3 ~ "High School (9-12) or GED",
    CAN_EDUCATION == 4 ~ "Attended College/Technical School",
    CAN_EDUCATION == 5 ~ "Associate/Bachelor Degree",
    CAN_EDUCATION == 6 ~ "Post-College Graduate Degree",
    CAN_EDUCATION == 996 ~ "N/A (< 5 years old)",
    CAN_EDUCATION == 998 ~ "Unknown",
    TRUE ~ "Other")
)
heart_adi_table = heart_SVM |> 
  select(CAN_AGE_AT_LISTING, Male, can_race, durable_LVAD,  initial_status, SVM_PERCENTILE, SVI_percentile, mean_ADI, dci_percentile, CAN_ETHNICITY_SRTR, status_table1, adi_decile, CAN_ABO, CAN_WGT_KG, CAN_HGT_CM, multi_listed, education, insurance) |> 
  filter(adi_decile == 1 | adi_decile == 10) |> 
  group_by(PX_ID) |> 
  tbl_summary(by = adi_decile, label = list(
    can_race ~ "Race",
    initial_status ~ "Initial Status",
    CAN_AGE_AT_LISTING ~ "Age",
    durable_LVAD ~ "Durable LVAD",
    SVM_PERCENTILE ~ "SVM Percentile",
    SVI_percentile ~ "SVI Percentile",
    mean_ADI ~ "ADI Percentile",
    dci_percentile ~ "DCI Percentile",
    CAN_ETHNICITY_SRTR ~ "Ethnicity",
    status_table1 ~ "Outcome",
    CAN_ABO ~ "Blood Type",
    CAN_WGT_KG ~ "Weight (kg)",
    CAN_HGT_CM ~ "Height (cm)",
    multi_listed ~ "Multi Organ Listing",
    education ~ "Education Level",
    insurance ~ "Primary Payment Type"
    
  ),
  statistic = list(all_continuous() ~ "{mean} ({sd})", all_categorical() ~ "{n} ({p}%)")) |> 
  modify_table_body(filter, !(variable == "PX_ID" )) 

heart_table = heart_SVM |> 
  select(CAN_AGE_AT_LISTING, Male, can_race, durable_LVAD,  initial_status, SVM_PERCENTILE, SVI_percentile, mean_ADI, dci_percentile, CAN_ETHNICITY_SRTR, status_table1, adi_decile, CAN_ABO, CAN_WGT_KG, CAN_HGT_CM, multi_listed, education, insurance) |> 
  group_by(PX_ID) |> 
  tbl_summary(label = list(
    can_race ~ "Race",
    initial_status ~ "Initial Status",
    CAN_AGE_AT_LISTING ~ "Age",
    durable_LVAD ~ "Durable LVAD",
    SVM_PERCENTILE ~ "SVM Percentile",
    SVI_percentile ~ "SVI Percentile",
    mean_ADI ~ "ADI Percentile",
    dci_percentile ~ "DCI Percentile",
    CAN_ETHNICITY_SRTR ~ "Ethnicity",
    status_table1 ~ "Outcome",
    CAN_ABO ~ "Blood Type",
    CAN_WGT_KG ~ "Weight (kg)",
    CAN_HGT_CM ~ "Height (cm)",
    multi_listed ~ "Multi Organ Listing",
    education ~ "Education Level",
    insurance ~ "Primary Payment Type"
    
  ),
  statistic = list(all_continuous() ~ "{mean} ({sd})", all_categorical() ~ "{n} ({p}%)")) |> 
  modify_table_body(filter, !(variable == "PX_ID" )) 

heart_SVM |> 
  select(CAN_AGE_AT_LISTING, CAN_GENDER, can_race, durable_LVAD,  initial_status, SVM_PERCENTILE, SVI_percentile, mean_ADI, dci_percentile, CAN_ETHNICITY_SRTR, status_table1, adi_decile, CAN_ABO, CAN_WGT_KG, CAN_HGT_CM, multi_listed, CAN_EDUCATION,  CAN_PRIMARY_PAY) |> 
  group_by(PX_ID) |> 
  tbl_summary(label = list(
    can_race ~ "Race",
    CAN_GENDER ~ "Sex"
  ),
  statistic = list(all_continuous() ~ "{mean} ({sd})", all_categorical() ~ "{n} ({p}%)"))

tbl_html <- as.data.frame(heart_adi_table)
html_table <- xtable(tbl_html)
library(xtable)
print(html_table, type = "html", file = "adi_heart_table.html")

heart_adi_table
```


#impute missing height and weight
```{r}
heart_SVM$CAN_HGT_CM[is.na(heart_SVM$CAN_HGT_CM)] <-median(heart_SVM$CAN_HGT_CM, na.rm=T)
heart_SVM$CAN_WGT_KG[is.na(heart_SVM$CAN_WGT_KG)] <-median(heart_SVM$CAN_WGT_KG, na.rm=T)



```


 
Fine-Gray CRR - unadjusted

```{r}
svm_un =  crr(Surv(survival_time,status) ~ SVM_decile, data = heart_SVM)

svm_death_un = crr(Surv(survival_time,status) ~ SVM_decile, data = heart_SVM, failcode = "death")


svi_un <- crr(Surv(survival_time,status) ~ SVI_decile, data = heart_SVM)

svi_death_un <- crr(Surv(survival_time,status) ~ SVI_decile , data = heart_SVM, failcode = "death")


adi_un <- crr(Surv(survival_time,status) ~ ADI_decile , data = heart_SVM)


adi_death_un <- crr(Surv(survival_time,status) ~ ADI_decile, data = heart_SVM, failcode = "death")


dci_un <- crr(Surv(survival_time,status) ~ DCI_decile, data = heart_SVM)


dci_death_un <- crr(Surv(survival_time,status) ~ DCI_decile, data = heart_SVM, failcode = "death")

dci_death_un

svm_tidy_un = svm_un$tidy
svi_tidy_un = svi_un$tidy
adi_tidy_un = adi_un$tidy
dci_tidy_un = dci_un$tidy
svm_tidy_death_un = svm_death_un$tidy
svi_tidy_death_un = svi_death_un$tidy
adi_tidy_death_un = adi_death_un$tidy
dci_tidy_death_un = dci_death_un$tidy
```

#make tables for model results
```{r}
create_clean_table <- function(data) {
  # Process the data into the clean table format
  clean_table <- data %>%
    mutate(
      Hazard_Ratio = exp(estimate), # Calculate Hazard Ratio
      Confidence_Interval = paste0("(", round(exp(conf.low), 2), ", ", round(exp(conf.high), 2), ")") # Create CI as a string
    ) %>%
    select(
      Variable = term,
      Coefficient = estimate,
      Hazard_Ratio,
      `Standard Error` = std.error,
      `P-Value` = p.value,
      `Confidence Interval` = Confidence_Interval
    )
  return(as.data.frame(clean_table))
}




```

#Partially adjusted models
```{r}

svm_small <- crr(Surv(survival_time,status) ~ SVM_decile + CAN_ABO + CAN_GENDER + CAN_HGT_CM + CAN_WGT_KG, data = heart_SVM)


svm_death_small <- crr(Surv(survival_time,status) ~ SVM_decile + CAN_ABO + CAN_GENDER + CAN_HGT_CM + CAN_WGT_KG, data = heart_SVM, failcode = "death")


svi_small <- crr(Surv(survival_time,status) ~ SVI_decile + CAN_ABO +  CAN_GENDER + CAN_HGT_CM + CAN_WGT_KG, data = heart_SVM)

svi_death_small <- crr(Surv(survival_time,status) ~ SVI_decile + CAN_ABO  + CAN_GENDER + CAN_HGT_CM + CAN_WGT_KG, data = heart_SVM, failcode = "death")


adi_small <- crr(Surv(survival_time,status) ~ ADI_decile + CAN_ABO + CAN_GENDER + CAN_HGT_CM + CAN_WGT_KG, data = heart_SVM)


adi_death_small <- crr(Surv(survival_time,status) ~ ADI_decile + CAN_ABO  + CAN_GENDER + CAN_HGT_CM + CAN_WGT_KG, data = heart_SVM, failcode = "death")


dci_small <- crr(Surv(survival_time,status) ~ DCI_decile + CAN_ABO +  CAN_GENDER + CAN_HGT_CM + CAN_WGT_KG, data = heart_SVM)


dci_death_small <- crr(Surv(survival_time,status) ~ DCI_decile + CAN_ABO + CAN_GENDER + CAN_HGT_CM + CAN_WGT_KG, data = heart_SVM, failcode = "death")

svm_tidy_small = svm_small$tidy
svi_tidy_small = svi_small$tidy
adi_tidy_small = adi_small$tidy
dci_tidy_small = dci_small$tidy
svm_tidy_death_small = svm_death_small$tidy
svi_tidy_death_small = svi_death_small$tidy
adi_tidy_death_small = adi_death_small$tidy
dci_tidy_death_small = dci_death_small$tidy
```
#Fully adjusted models
```{r}

svm_full <- crr(Surv(survival_time,status) ~ SVM_decile + CAN_ABO +  CAN_GENDER + CAN_HGT_CM + CAN_WGT_KG + initial_status + durable_LVAD + multi_listed, data = heart_SVM)


svm_death_full <- crr(Surv(survival_time,status) ~ SVM_decile + CAN_ABO + CAN_GENDER + CAN_HGT_CM + CAN_WGT_KG + initial_status + durable_LVAD + multi_listed, data = heart_SVM, failcode = "death")


svi_full <- crr(Surv(survival_time,status) ~ SVI_decile + CAN_ABO + CAN_GENDER + CAN_HGT_CM + CAN_WGT_KG + initial_status + durable_LVAD + multi_listed, data = heart_SVM)

svi_death_full <- crr(Surv(survival_time,status) ~ SVI_decile + CAN_ABO + CAN_GENDER + CAN_HGT_CM + CAN_WGT_KG + initial_status + durable_LVAD + multi_listed, data = heart_SVM, failcode = "death")


adi_full <- crr(Surv(survival_time,status) ~ ADI_decile + CAN_ABO + CAN_GENDER + CAN_HGT_CM + CAN_WGT_KG + initial_status + durable_LVAD + multi_listed, data = heart_SVM)


adi_death_full <- crr(Surv(survival_time,status) ~ ADI_decile + CAN_ABO  + CAN_GENDER + CAN_HGT_CM + CAN_WGT_KG + initial_status + durable_LVAD + multi_listed, data = heart_SVM, failcode = "death")


dci_full <- crr(Surv(survival_time,status) ~ DCI_decile + CAN_ABO +  CAN_GENDER + CAN_HGT_CM + CAN_WGT_KG + initial_status + durable_LVAD + multi_listed, data = heart_SVM)


dci_death_full <- crr(Surv(survival_time,status) ~ DCI_decile + CAN_ABO +  CAN_GENDER + CAN_HGT_CM + CAN_WGT_KG + initial_status + durable_LVAD + multi_listed, data = heart_SVM, failcode = "death")

svm_tidy_full = svm_full$tidy
svi_tidy_full = svi_full$tidy
adi_tidy_full = adi_full$tidy
dci_tidy_full = dci_full$tidy
svm_tidy_death_full = svm_death_full$tidy
svi_tidy_death_full = svi_death_full$tidy
adi_tidy_death_full = adi_death_full$tidy
dci_tidy_death_full = dci_death_full$tidy
```
#save model results for supplement
#Supplemental Tables 1-24
```{r}

save_df_as_html <- function(data, file_name) {
  # Convert the data frame to an HTML table
  data = create_clean_table(data)
  html_table <- xtable(data)
  
  # Save the table as an HTML file
  print(html_table, type = "html", file = file_name)
  
  # Inform the user
  message("HTML table saved as ", file_name)
}

save_df_as_html(svm_tidy_un, "svm_table_un.html") # Supplemental Table 1
save_df_as_html(svi_tidy_un, "svi_table_un.html") # Supplemental Table 2
save_df_as_html(adi_tidy_un, "adi_table_un.html") # Supplemental Table 3
save_df_as_html(dci_tidy_un, "dci_table_un.html") # Supplemental Table 4
save_df_as_html(svm_tidy_death_un, "svm_table_death_un.html") # Supplemental Table 5
save_df_as_html(svi_tidy_death_un, "svi_table_death_un.html") # Supplemental Table 6
save_df_as_html(adi_tidy_death_un, "adi_table_death_un.html") # Supplemental Table 7
save_df_as_html(dci_tidy_death_un, "dci_table_death_un.html") # Supplemental Table 8

save_df_as_html(svm_tidy_small, "svm_table_small.html") # Supplemental Table 9
save_df_as_html(svi_tidy_small, "svi_table_small.html") # Supplemental Table 10
save_df_as_html(adi_tidy_small, "adi_table_small.html") # Supplemental Table 11
save_df_as_html(dci_tidy_small, "dci_table_small.html") # Supplemental Table 12
save_df_as_html(svm_tidy_death_small, "svm_table_death_small.html") # Supplemental Table 13
save_df_as_html(svi_tidy_death_small, "svi_table_death_small.html") # Supplemental Table 14
save_df_as_html(adi_tidy_death_small, "adi_table_death_small.html") # Supplemental Table 15
save_df_as_html(dci_tidy_death_small, "dci_table_death_small.html") # Supplemental Table 16

save_df_as_html(svm_tidy_full, "svm_table_full.html") # Supplemental Table 17
save_df_as_html(svi_tidy_full, "svi_table_full.html") # Supplemental Table 18
save_df_as_html(adi_tidy_full, "adi_table_full.html") # Supplemental Table 19
save_df_as_html(dci_tidy_full, "dci_table_full.html") # Supplemental Table 20
save_df_as_html(svm_tidy_death_full, "svm_table_death_full.html") # Supplemental Table 21
save_df_as_html(svi_tidy_death_full, "svi_table_death_full.html") # Supplemental Table 22
save_df_as_html(adi_tidy_death_full, "adi_table_death_full.html") # Supplemental Table 23
save_df_as_html(dci_tidy_death_full, "dci_table_death_full.html") # Supplemental Table 24
svm_tidy_death_un
```

## Figure 2: Cumulative Incidence Curves

```{r cumic curves}
heart_SVM$statuscuminc <- factor(heart_SVM$status_numeric,
                           levels = c(0,1,2,3),
                           labels = c("On waitlist at end of follow-up","Received transplant","Died", "Removed for other reason"))

# SVM Plot
svm_cuminc <- cuminc(Surv(survival_time, statuscuminc) ~ SVM_decile, data = subset(heart_SVM, (SVM_decile == "1st Decile" | SVM_decile == "10th Decile"))) %>% 
  ggcuminc(outcome = c("Received transplant", "Died")) +
  ylim(0, 1) +
  labs(
    x = "Days",
    title = "SVM"
  ) + 
  add_confidence_interval() +
  theme(
    plot.title = element_text(size = 18, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)
  )

# SVI Plot
svi_cuminc <- cuminc(Surv(survival_time, statuscuminc) ~ SVI_decile, data = subset(heart_SVM, (SVI_decile == "1st Decile" | SVI_decile == "10th Decile"))) %>% 
  ggcuminc(outcome = c("Received transplant", "Died")) +
  ylim(0, 1) +
  labs(
    x = "Days",
    title = "SVI"
  ) + 
  add_confidence_interval() +
  theme(
    plot.title = element_text(size = 18, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)
  )

# ADI Plot
adi_cuminc <- cuminc(Surv(survival_time, statuscuminc) ~ ADI_decile, data = subset(heart_SVM, (ADI_decile == "1st Decile" | ADI_decile == "10th Decile"))) %>% 
  ggcuminc(outcome = c("Received transplant", "Died")) +
  ylim(0, 1) +
  labs(
    x = "Days",
    title = "ADI"
  ) + 
  add_confidence_interval() +
  theme(
    plot.title = element_text(size = 18, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)
  )

# DCI Plot
dci_cuminc <- cuminc(Surv(survival_time, statuscuminc) ~ DCI_decile, data = subset(heart_SVM, (DCI_decile == "1st Decile" | DCI_decile == "10th Decile"))) %>% 
  ggcuminc(outcome = c("Received transplant", "Died")) +
  ylim(0, 1) +
  labs(
    x = "Days",
    title = "DCI"
  ) + 
  add_confidence_interval() +
  theme(
    plot.title = element_text(size = 18, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)
  )

# Arrange the plots in a grid
grid.arrange(svm_cuminc, svi_cuminc, adi_cuminc, dci_cuminc, ncol = 2, nrow = 2)
ggsave(
  filename = "cumincs_heart.png",
  plot = grid.arrange(svm_cuminc, svi_cuminc, adi_cuminc, dci_cuminc, ncol = 2, nrow = 2),
  width = 12, height = 10  # Adjust width and height as needed
)
svm_cuminc

tidy(cuminc(Surv(survival_time,statuscuminc) ~ ADI_decile, data = subset(heart_SVM, (ADI_decile == "1st Decile" | ADI_decile == "10th Decile")))) %>% filter(time == 1095)

adi_cuminc
```

#SVM forest plots unadjusted model
```{r svm forest}

forest <- tidy(svm_un)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)
forest

svm_forest_un <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  xlim(0.5,1.2)+
  labs(x = "",
       y = "SVM Decile",
       title = "Unadjusted Model")+
  theme_classic()
-0

forest <- tidy(svm_death_un)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)

svm_death_forest_un <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  labs(x = "",
       y = "SVM Decile",
       title = "Unadjusted Model")+
   xlim(0.25, 2)+
  theme_classic()

svm_forest_un
```
#SVI forest plots unadjusted model
```{r svi forest}

forest <- tidy(svi_un)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)
forest

svi_forest_un <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  xlim(0.5,1.2)+
  labs(x = "",
       y = "SVI Decile",
       title = "Unadjusted Model")+
  theme_classic()

forest <- tidy(svi_death_un)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)

svi_death_forest_un <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  labs(x = "",
       y = "SVI Decile",
       title = "Unadjusted Model")+
   xlim(0.15, 2)+
  theme_classic()
forest
```
#ADI forest plots unadjusted model

```{r adi forest}

forest <- tidy(adi_un)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)
forest
adi_forest_un <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point(color = "black") +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
    xlim(0.5,1.2)+
  labs(x = "",
       y = "ADI Decile",
       title = "Unadjusted Model")+
  theme_classic()

forest <- tidy(adi_death_un)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)
forest
adi_death_forest_un <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point(color = "black") +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  labs(x = "",
       y = "ADI Decile",
       title = "Unadjusted Model")+
  xlim(0.25, 5)+
  theme_classic()

```

#DCI forest plots unadjusted model
```{r dci forest}

forest <- tidy(dci_un)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)
forest

dci_forest_un <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
    xlim(0.5,1.2)+
  labs(x = "",
       y = "DCI Decile",
       title = "Unadjusted Model")+
  theme_classic()

forest <- tidy(dci_death_un)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)

dci_death_forest_un <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  labs(x= "",
       y = "DCI Decile",
       title = "Unadjusted Model")+
  xlim(0.25, 2)+
  theme_classic()
```




#SVM forest plots partially adjusted model

```{r svm forest}

forest <- tidy(svm_small)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)
forest
svm_forest_small <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  xlim(0.5,1.2)+
  labs(x = "Sub-Hazard Ratio of Transplant",
       y = "SVM Decile",
       title = "Partially adjusted")+
  theme_classic()


forest <- tidy(svm_death_small)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)

svm_death_forest_small <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  labs(x = "Sub-Hazard Ratio of Death",
       y = "SVM Decile",
       title = "Partially adjusted")+
   xlim(0.25, 2)+
  theme_classic()
```
#SVI forest plots partially adjusted model
```{r svi forest}

forest <- tidy(svi_small)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)
forest

svi_forest_small <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  xlim(0.5,1.2)+
  labs(x = "Sub-Hazard Ratio of Transplant",
       y = "SVI Decile",
       title = "Partially adjusted")+
  theme_classic()

forest <- tidy(svi_death_small)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)

svi_death_forest_small <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  labs(x = "Sub-Hazard Ratio of Death",
       y = "SVI Decile",
       title = "Partially adjusted")+
   xlim(0.15, 2)+
  theme_classic()
```

#ADI forest plots partially adjusted model
```{r adi forest}

forest <- tidy(adi_small)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)
forest
adi_forest_small <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point(color = "black") +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
    xlim(0.5,1.2)+
  labs(x = "Sub-Hazard Ratio of Transplant",
       y = "ADI Decile",
       title = "Partially adjusted")+
  theme_classic()

forest <- tidy(adi_death_small)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)
forest
adi_death_forest_small <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point(color = "black") +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  labs(x = "Sub-Hazard Ratio of Death",
       y = "ADI Decile",
       title = "Partially adjusted")+
  xlim(0.25, 5)+
  theme_classic()

```

#DCI forest plots partially adjusted model
```{r dci forest}

forest <- tidy(dci_small)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)
forest
dci_forest_small <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
    xlim(0.5,1.2)+
  labs(x = "Sub-Hazard Ratio of Transplant",
       y = "DCI Decile",
       title = "Partially adjusted")+
  theme_classic()

forest <- tidy(dci_death_small)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)

dci_death_forest_small <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  labs(x = "Sub-Hazard Ratio of Death",
       y = "DCI Decile",
       title = "Partially adjusted")+
  xlim(0.25, 2)+
  theme_classic()
```


#SVM forest plots full model
```{r svm forest}

forest <- tidy(svm_full)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)
forest

svm_forest_full <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  xlim(0.5,1.2)+
  labs(x = "",
       y = "SVM Decile",
       title = "Full model")+
  theme_classic()


forest <- tidy(svm_death_full)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)

svm_death_forest_full <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  labs(x = "",
       y = "SVM Decile",
       title = "Full model")+
   xlim(0.25, 2)+
  theme_classic()
```
#SVI forest plots full model
```{r svi forest}

forest <- tidy(svi_full)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)
forest

svi_forest_full <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  xlim(0.5,1.2)+
  labs(x = "",
       y = "SVI Decile",
       title = "Full model")+
  theme_classic()

forest <- tidy(svi_death_full)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)

svi_death_forest_full <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  labs(x = "",
       y = "SVI Decile",
       title = "Full model")+
   xlim(0.15, 2)+
  theme_classic()
```
#ADI forest plots full model
```{r adi forest}

forest <- tidy(adi_full)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)
forest
adi_forest_full <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point(color = "black") +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
    xlim(0.5,1.2)+
  labs(x = "",
       y = "ADI Decile",
       title = "Full model")+
  theme_classic()

forest <- tidy(adi_death_full)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)
forest
adi_death_forest_full <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point(color = "black") +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  labs(x = "",
       y = "ADI Decile",
       title = "Full model")+
  xlim(0.25, 5)+
  theme_classic()
forest
adi_death_forest_full
```
#DCI forest plots full model
```{r dci forest}

forest <- tidy(dci_full)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)
forest

dci_forest_full <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
    xlim(0.5,1.2)+
  labs(x = "",
       y = "DCI Decile",
       title = "Full model")+
  theme_classic()

forest <- tidy(dci_death_full)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)

dci_death_forest_full <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  labs(x = "",
       y = "DCI Decile",
       title = "Full model")+
  xlim(0.25, 2)+
  theme_classic()
```
# Supplemental Figure 2: SVM Transplant Forest Plot
# Supplemental Figure 5: SVM Death Forest Plot
```{r}
grid.arrange(svm_forest_un,svm_forest_small, svm_forest_full, ncol =3, nrow = 1)

grid.arrange(svm_death_forest_un, svm_death_forest_small, svm_death_forest_full, ncol=3, nrow = 1)


grid_tx <- arrangeGrob(svm_forest_un,svm_forest_small, svm_forest_full, ncol =3, nrow = 1)  # Arrange in a 2-column grid
ggsave("svm_transplant.png", grid_tx, width = 12, height = 7)
grid_death <- arrangeGrob(svm_death_forest_un, svm_death_forest_small, svm_death_forest_full, ncol=3, nrow = 1)
ggsave("svm_death.png", grid_death, width = 12, height = 7)
```
# Supplemental Figure 3: SVI Transplant Forest Plot
# Supplemental Figure 6: SVI Death Forest Plot
```{r figure 3}
grid.arrange(svi_forest_un,svi_forest_small, svi_forest_full, ncol =3, nrow = 1)
grid.arrange(svi_death_forest_un, svi_death_forest_small, svi_death_forest_full, ncol=3, nrow = 1)

grid_tx <- arrangeGrob(svi_forest_un,svi_forest_small, svi_forest_full, ncol =3, nrow = 1)  # Arrange in a 2-column grid
ggsave("svi_transplant.png", grid_tx, width = 12, height = 7)
grid_death <- arrangeGrob(svi_death_forest_un, svi_death_forest_small, svi_death_forest_full, ncol=3, nrow = 1)
ggsave("svi_death.png", grid_death, width = 12, height = 7)

```
# Figure 3: ADI Transplant Forest Plot
# Figure 4: ADI Death Forest Plot
```{r figure 3}
grid_tx <- arrangeGrob(adi_forest_un,adi_forest_small, adi_forest_full, ncol =3, nrow = 1)  # Arrange in a 2-column grid
ggsave("adi_transplant.png", grid_tx, width = 12, height = 7)
grid_death <- arrangeGrob(adi_death_forest_un,adi_death_forest_small, adi_death_forest_full, ncol =3, nrow = 1)
ggsave("adi_death.png", grid_death, width = 12, height = 7)
```

# Supplemental Figure 4: DCI Transplant Forest Plot
# Supplemental Figure 7: DCI Death Forest Plot
```{r figure 3}
grid.arrange(dci_forest_un,dci_forest_small, dci_forest_full, ncol =3, nrow = 1)

grid.arrange(dci_death_forest_un, dci_death_forest_small, dci_death_forest_full, ncol=3, nrow = 1)

grid_tx <- arrangeGrob(dci_forest_un,dci_forest_small, dci_forest_full, ncol =3, nrow = 1)  # Arrange in a 2-column grid
ggsave("dci_transplant.png", grid_tx, width = 12, height = 7)
grid_death <- arrangeGrob(dci_death_forest_un, dci_death_forest_small, dci_death_forest_full, ncol=3, nrow = 1)
ggsave("dci_death.png", grid_death, width = 12, height = 7)
```
# Durable LVAD usage by Decile
```{r}
lvad_svm <- heart_SVM %>%
  group_by(svm_decile, durable_LVAD) %>%   # Group by decile and durable_LVAD
  dplyr::summarize(count = n(), .groups = "drop") %>% # Count occurrences within each group
  as.data.frame() %>%  # Convert to a data frame (optional)
  group_by(svm_decile) %>%  # Group by decile only to calculate percentages
  mutate(SVM = (count / sum(count)) * 100) %>%  # Calculate percentage for each decile
  ungroup() %>%  # Remove grouping
  filter(durable_LVAD == 1) %>%  # Keep only rows where durable_LVAD is 1
  select(svm_decile, SVM) 
lvad_svm

## SVI
lvad_svi <- heart_SVM %>%
  group_by(svi_decile, durable_LVAD) %>% # Group by decile and blood type
  dplyr::summarize(count = n(), .groups = "drop") %>% # Create count column
  as.data.frame() %>% # Convert to data frame
  group_by(svi_decile) %>% # Group by decile
  mutate(SVI = (count/sum(count))*100) %>% # Create percentage column
  ungroup() %>% 
  filter(durable_LVAD == 1) %>% # Keep only blood type O
  select(svi_decile, SVI) # Keep columns of interest
lvad_svi

## ADI
lvad_adi <- heart_SVM %>%
  group_by(adi_decile, durable_LVAD) %>% # Group by decile and blood type
  dplyr::summarize(count = n(), .groups = "drop") %>% # Create count column
  as.data.frame() %>% # Convert to data frame
  group_by(adi_decile) %>% # Group by decile
  mutate(ADI = (count/sum(count))*100) %>% # Create percentage column
  ungroup() %>% 
  filter(durable_LVAD == 1) %>% # Keep only blood type O
  select(adi_decile, ADI) # Keep columns of interest
lvad_adi

## DCI
lvad_dci <- heart_SVM %>%
  group_by(dci_decile, durable_LVAD) %>% # Group by decile and blood type
  dplyr::summarize(count = n(), .groups = "drop") %>% # Create count column
  as.data.frame() %>% # Convert to data frame
  group_by(dci_decile) %>% # Group by decile
  mutate(DCI = (count/sum(count))*100) %>% # Create percentage column
  ungroup() %>% 
  filter(durable_LVAD == 1) %>% # Keep only blood type O
  select(dci_decile, DCI) # Keep columns of interest
lvad_dci = head(lvad_dci, 10)

lvad_svi

```

```{r}
lvad_dci
lvad_df <- cbind(lvad_svm, lvad_svi, lvad_adi, lvad_dci) %>% # Join data sets by columns
  select(-svi_decile, -adi_decile, dci_decile) %>% # Remove extra decile columns
  rename(decile = svm_decile) # Change name to "decile"
lvad_df

rm(lvad_adi, lvad_dci, lvad_svi, lvad_svm) # Remove intermediate data

# Reshape data set
lvad_df <- lvad_df %>%
  pivot_longer(cols = c(SVM, SVI, ADI, DCI), # Select columns to reshape data by
               names_to = "index", # Set names to index column
               values_to = "percent") # Set values to percent column
lvad_df

# Bar graph w/o facets (Figure 5: Durable LVAD usage by Decile)
lvad_df_visual <- 
  ggplot(lvad_df, aes(x = decile, y = percent, fill = index)) + # Set base
  geom_col(position = "dodge") + # Set columns side-by-side
  labs(title = "Percentage of Durable LVAD by Decile", # Set title
       subtitle = "Data from SRTR 2024", # Set subtitle
       x = "Decile", # Set x-axis label
       y = "Percentage",  # Set y-axis label
       fill = "Index") + # Set legend title
  theme_classic() + # Set theme
  scale_y_continuous(limits = c(0, 40),
                     breaks = seq(0, 40, by = 10)) +
  scale_x_continuous(breaks = seq(1,10))# Set y-axis 
lvad_df_visual

ggsave("lvad_deciles.png", lvad_df_visual, width = 10, height = 7)

# Bar graph w/ facets
facet_lvad_df_visual <- 
  ggplot(lvad_df, aes(x = decile, y = percent, fill = index)) + # Set base
  geom_col(position = "dodge") + # Set columns side-by-side
  facet_wrap(~ index) + # Create facet grid
  labs(title = "Percentage of Durable LVAD by Decile", # Set title
       subtitle = "Data from SRTR 2024", # Set subtitle
       x = "Decile",  # Set x-axis label
       y = "Percentage", # Set y-axis label
       fill = "Index") + # Set legend title
  theme_bw() + # Set theme
  scale_y_continuous(limits = c(0, 60),
                     breaks = seq(0, 60, by = 10)) +
  scale_x_continuous(breaks = seq(1,10)) # Set y-axis
facet_lvad_df_visual


```


