---
title: "Compare Deprivation Index Heart"
output: html_document
date: "2024-05-23"
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(dplyr)
library(survival)
library(haven)
library(lubridate)
library(ggsurvfit)
library(gtsummary)
library(readr)
library(consort)
library(splines)
library(readxl)
library(rms)
library(forcats)
library(cmprsk)
library(crrSC)
library(sf)
library(forestplot)
library(gridExtra)
library(pROC)
library(coxme)
```

## Loading Data

```{r data}
setwd("~/R/Parker")
cand_thor <- read_sas("cand_thor.sas7bdat", NULL)
slice_head(cand_thor,n=5)
```
## ASsigning Scores to Zip codes

## ADI

ADI is coded to the at the block level (12 digit FIPS), whereas the crosswalk document is only coded at the tract level (11 digit FIPS). Need to adjust accordingly. ADI from University of Wisconsin. Crosswalk document from US Department of Housing and Urban Dev.

```{r}
ADI <- read_csv("2021_ADI.csv")
ADI_crosswalk <-read_excel("11_digit_FIPS_to_ZIP_crosswalk.xlsx")

ADI$FIPS_11_digit <- substr(ADI$FIPS, 1, 11) 
ADI <- left_join(ADI,
                 select(ADI_crosswalk, TRACT, ZIP), 
                 by = c("FIPS_11_digit" = "TRACT") )
ADI <- ADI |> 
  group_by(ZIP) |> 
  mutate(mean_ADI = mean(as.numeric(ADI_NATRANK), na.rm = TRUE))
ADI |> 
  select(ZIP,  ADI_NATRANK, mean_ADI) |> 
  slice_head(n = 10)
ADI <- distinct(ADI, ZIP, .keep_all = TRUE)
rm(ADI_crosswalk)

```

## SVI

SVI is stupid and codes missing data as -999. 

```{r}
SVI <- read_csv("SVI_2022_US.csv")
SVI_crosswalk <- read_excel("11_digit_FIPS_to_ZIP_crosswalk.xlsx")

SVI <- left_join(SVI, select(SVI_crosswalk, TRACT, ZIP), by = c("FIPS" = "TRACT"))
SVI <- SVI |> 
  select(ZIP, RPL_THEMES) 
SVI <- SVI |> mutate(
  RPL_THEMES = case_when(RPL_THEMES < 0 ~ NA, .default = RPL_THEMES))
SVI <- SVI |> 
  group_by(ZIP) |> 
  mutate(SVI_percentile = mean(as.numeric(RPL_THEMES*100), na.rm = TRUE),
         median_SVI = median(as.numeric(RPL_THEMES*100), na.rm = TRUE))

SVI |>
  group_by(ZIP) |> 
  ggplot() +
  geom_point(mapping = aes(x = SVI_percentile, y = median_SVI))
SVI <- distinct(SVI, ZIP, .keep_all = TRUE)
rm(SVI_crosswalk)
slice_head(SVI, n=10) 


```

## SVM

```{r}
SVM_ZIP <- read_csv("SVM_ZIPCODE_SCORES_All States_All Counties_2024-01-23.csv")
SVM_ZIP$zipcode <- substr(SVM_ZIP$ZIPCODE, start = 1, stop = 5)
slice_head(SVM_ZIP,n= 5)

```
## DCI

```{r}
DCI <- read_excel("DCI-2017-2021-Scores-Only.xlsx")
DCI <- DCI |> 
  mutate(zip = as.character(`Zip Code`),
         dci_percentile = `2017-2021 Final Distress Score`)
slice_head(DCI, n = 5)
```


## Assign Deprivation Inidicies to Candidate ZIP Code

Good to know how many people have the 9 digit zip code

```{r}
cand_zip <- read_sas("canzip2403.sas7bdat", NULL)
cand_zip$ZIP = substr(cand_zip$CAN_PERM_ZIP,1,5)
cand_zip <- left_join(cand_zip, 
                              select(SVM_ZIP, ZIPCODE, SVM_SCORE, SVM_PERCENTILE),
                              by = c("ZIP" = "ZIPCODE"))
cand_zip <-left_join(cand_zip, SVI, by = c("ZIP" = "ZIP"))
cand_zip <- left_join(cand_zip, select(ADI, mean_ADI, ZIP), by = c("ZIP" = "ZIP"))
cand_zip <- left_join(cand_zip, select(DCI, dci_percentile, zip), by = c("ZIP" = "zip"))
rm(list = c("ADI", "SVI", "SVM_ZIP", "DCI"))

slice_head(cand_zip,n= 5)
```

## Subet Cohort

Only include patients added to the waitlist through 2019 or 2022.This allows us to observe everyone for at least 1 year this way. Only patients that are excluded are patients that are simultaneously listed for heart and lung.

```{r}
cand_thor <- cand_thor |> 
  subset(CAN_LISTING_DT > as.Date("2019-01-01") & CAN_LISTING_DT < as.Date("2022-12-31") & WL_ORG == "HR")
```

## VAD Type characterization

Credit to Kevin L. (https://github.com/kevinlazenby/post_transplant_survival_new_policy/blob/main/function_definitions.Rmd), lines 427 - 460

Durable LVADs (line 867) https://github.com/healthallocate/Heart-Continuous-Score/blob/main/6week_death_data_prep.Rmd


```{r}
cand_thor <- cand_thor |> 
  mutate(VAD = case_when(
        CAN_VAD_TY == 1 ~ paste0("None"),
        CAN_VAD_TY == 2 ~ paste0("LVAD"),
        CAN_VAD_TY == 3 ~ paste0("RVAD"),
        CAN_VAD_TY == 4 ~ paste0("TAH"),
        CAN_VAD_TY == 5 ~ paste0("LVAD+RVAD")),
      
      LVAD = case_when(
        VAD == "LVAD" ~ 1,
        TRUE ~ 0),
      
      durable_LVAD = case_when(
        CAN_VAD1 %in% c('202', '205', '206', '207', '208', '209', '210', 
                        '212', '213', '214', '223', '224', '233', '236', 
                        '239', '240', '312', '313', '314', '315', '316', '319', '322', 
                        '327', '330', '333', '334') | 
          CAN_VAD2 %in%  c('202', '205', '206', '207', '208', '209', '210', 
                        '212', '213', '214', '223', '224', '233', '236', 
                        '239', '240', '312', '313', '314', '315', '316', '319', '322', 
                        '327', '330', '333', '334') ~ 1,
        TRUE ~ 0),

      
      temporary_LVAD = case_when(
        durable_LVAD != 1 & !is.na(CAN_VAD1) ~ 1,
        TRUE ~ 0),
      
      RVAD = case_when(
        VAD == "RVAD" ~ 1,
        (CAN_VAD_TY != 1 & !is.na(CAN_VAD_TY) & 
       !(CAN_VAD1 %in% c('205', '236', '313', '330', '206', '208', '314', 
                         '210', '319', '216', '305', '217', '306', '223', 
                         '312', '224', '316', '230', '324', '231', '325', 
                         '232', '326', '233', '327'))) ~ 1,
        TRUE ~ 0),
      
      TAH = case_when(
        VAD == "TAH" ~ 1,
        TRUE ~ 0),
      
      LVAD_RVAD = case_when(
        VAD == "LVAD+RVAD" ~ 1,
        TRUE ~ 0))
  
      
```


## Intervals

Credit to Kevin Z. in USCRS score (lines 511 - 538)
https://github.com/healthallocate/Heart-Continuous-Score/blob/main/6week_death_data_prep.Rmd


```{r intervals}
 
cand_thor <- cand_thor |> 
  mutate(
    transplant_cand_thor = case_when(!is.na(REC_TX_DT) | CAN_REM_CD == '4' ~ 1,
      TRUE ~ 0),
    start_date = as.Date(CAN_LISTING_DT),
    transplant_date = as.Date(REC_TX_DT),
    removal_date = as.Date(CAN_REM_DT),
    last_active_date = as.Date(CAN_LAST_ACT_STAT_DT),
    last_inactive_date = as.Date(CAN_LAST_INACT_STAT_DT),
    death_date = as.Date(CAN_DEATH_DT),
    death_date_max = pmax(CAN_DEATH_DT, PERS_SSA_DEATH_DT, PERS_OPTN_DEATH_DT, na.rm=T)) %>%
  
  mutate(transplant_time = ifelse(!is.na(transplant_date), 
                                  as.numeric(transplant_date - start_date, units='days'), NA)) %>%
  mutate(death_time = ifelse(!is.na(death_date_max), as.numeric(death_date_max - start_date, units='days'), NA)) %>%
  mutate(removal_time = as.numeric(pmax(removal_date, last_active_date, last_inactive_date, na.rm = T) - 
                                     start_date, units='days'))
```

## Survival time

Survival time defined as time to transplant, death, or removal from waitlist, whichever comes first. Units = days (from previous code)

```{r survival time}
 cand_thor <- 
  cand_thor |> 
  mutate(
    survival_time = pmin(transplant_time, death_time, removal_time, na.rm = TRUE),
    survival_time = case_when(
      survival_time > 1095 ~ 1095,
      .default = survival_time
    ))
```

## Censoring

Death or deterioration on the waitlist coded as the event (1). Transplant is the censoring event (0).

```{r censoring}
cand_thor <- cand_thor |> 
  group_by(PX_ID) |> 
    mutate(status = case_when(
      survival_time == 1095 ~ 0, 
      survival_time == death_time ~ 2,
      CAN_REM_CD == 13 ~ 2,
      CAN_REM_CD %in% c(4, 14, 15, 18, 19, 20, 22, 23)  ~ 1,
      .default = 0
      ))

cand_thor$status <- factor(cand_thor$status,
                           levels = c(0,1,2),
                           labels = c("censor",  "transplant","death/deteriorate"))
```

Cause Specific Hazard setup

```{r}
cand_thor <- cand_thor |> 
  mutate(transplant_outcome = case_when(
    status == "censor" ~ 0,
    status == "transplant" ~ 1,
    .default = 0),
    death_outcome = case_when(
      status == "censor" ~ 0,
      status == "death/deteriorate" ~ 1,
      .default = 0
    ))
```

## STROBE/CONSORT Diagram

Inclusion and exclusion criteria Strobe diagram

```{r}
cand_thor <- cand_thor |> 
  mutate(eligibility = case_when(
    CAN_AGE_AT_LISTING < 18 ~ "Pediatric candidates",
    CAN_LISTING_DT == CAN_DEATH_DT ~ "Candidate listed and died same day",
  .default = NA),
    durable_LVAD_table1 = case_when(
        (is.na(eligibility) & durable_LVAD == 1) ~ "Durable LVAD",
        (is.na(eligibility) & durable_LVAD == 0) ~ "No Durable LVAD",
        .default = NA
      ),
  status_table1 = case_when(
    status =="censor" ~ "On waitlist at end of follow-up",
    status == "transplant" ~ "Received transplant",
    status == 'death/deteriorate' ~ "Died or removed from waitlist for deterioration"
  ))



txt <- gen_text(cand_thor$PX_ID, label = "Heart transplant candidate registrations, 2019-2022")
p_cons <- add_box(txt = txt)

txt <- gen_text(cand_thor$eligibility, label = "Excluded", bullet = TRUE)
p_cons <- add_side_box(p_cons, txt = txt)

p_cons <- add_box(p_cons, txt = gen_text(cand_thor$durable_LVAD_table1, label = "Adult candidate registrations"))

## Creating Analytic Sample 
cand_thor <- cand_thor |> 
  subset(is.na(eligibility))

txt <- gen_text(cand_thor$status_table1, bullet = TRUE)
p_cons <- add_box(p_cons, txt = txt, just = "left")


plot(p_cons, grViz = TRUE)
```


## Assigning Deprivation Indices based on Patient's permanent zip

Change to Person area code eventually

```{r center scores}
heart_SVM <- left_join(cand_thor,
                    select(cand_zip,SVM_SCORE, SVM_PERCENTILE, PX_ID, SVI_percentile, mean_ADI, dci_percentile), 
                    by = c("PX_ID" = "PX_ID"))
```

## Splitting SVM into Decile

```{r}
heart_SVM <- heart_SVM |> 
  group_by(PX_ID) |> 
  mutate(SVM_decile = 
           cut(SVM_PERCENTILE, c(-Inf, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100), labels = c("1st Decile", "2nd Decile", "3rd Decile", "4th Decile", "5th Decile", "6th Decile", "7th Decile", "8th Decile", "9th Decile", "10th Decile")),
           SVI_decile = cut(SVI_percentile, c(-Inf, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100), labels = c("1st Decile", "2nd Decile", "3rd Decile", "4th Decile", "5th Decile", "6th Decile", "7th Decile", "8th Decile", "9th Decile", "10th Decile")),
           ADI_decile = cut(mean_ADI,c(-Inf, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100), labels = c("1st Decile", "2nd Decile", "3rd Decile", "4th Decile", "5th Decile", "6th Decile", "7th Decile", "8th Decile", "9th Decile", "10th Decile")),
         DCI_decile = cut(dci_percentile,c(-Inf, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100), labels = c("1st Decile", "2nd Decile", "3rd Decile", "4th Decile", "5th Decile", "6th Decile", "7th Decile", "8th Decile", "9th Decile", "10th Decile")))
heart_SVM <- heart_SVM |> 
  mutate(svm_decile = as.numeric(SVM_decile),
        svi_decile = as.numeric(SVI_decile),
        adi_decile = as.numeric(ADI_decile),
        dci_decile = as.numeric(DCI_decile))
        
```



## Covariates

```{r covariates}
heart_SVM <- heart_SVM |> 
  mutate(can_race = case_when(
    CAN_RACE == "8" ~ "White",
    CAN_RACE == "16" ~ "Black or African American",
    CAN_RACE == "32" ~ "American Indian or Alaska Native",
    CAN_RACE == "64" ~ "Asian",
    CAN_RACE == "128" ~ "Native Hawaiian or Other Pacific Islander",
    CAN_RACE == "256" ~ "Arab or Middle Eastern",
    .default = "Other"),
    can_race = as.factor(can_race),
    ethnicity = case_when(
      CAN_ETHNICITY_SRTR == "NLATIN" ~ "Not Latinx",
      CAN_ETHNICITY_SRTR == "LATINO" ~ "Latinx"
      
    ),
    ethnicity = as.factor(ethnicity))
heart_SVM$can_race <- relevel(heart_SVM$can_race, ref = "White")
heart_SVM$ethnicity <- relevel(heart_SVM$ethnicity, ref = "Not Latinx")
heart_SVM <- heart_SVM |> 
  mutate(initial_status = case_when(
           CAN_INIT_STAT %in% c('1010', '1020', '1090', '1110', '2010', '2020', '2090', '2110', '3010', '6010', '6011', '6012', '9010') ~ "Status 1",
           CAN_INIT_STAT %in% c('1030', '1120', '2030', '2120', '6002', '6020', '6030', '9020') ~ "Status 2",
           CAN_INIT_STAT %in% c('1130', '2130', '6040') ~ "Status 3",
           CAN_INIT_STAT %in% c('1140', '2140', '6004') ~ "Status 4",
           CAN_INIT_STAT %in% c('1150', '2150') ~ "Status 5",
           CAN_INIT_STAT %in% c('1160', '2160') ~ "Status 6",
           .default = NA
         ))
```



## Table 1

```{r table 1}
heart_SVM |> 
  select(CAN_AGE_AT_LISTING, CAN_GENDER, can_race, durable_LVAD, temporary_LVAD, RVAD, LVAD_RVAD, TAH, initial_status, SVM_PERCENTILE, SVI_percentile, mean_ADI, dci_percentile, status_table1, CAN_ETHNICITY_SRTR, adi_decile ) |> 
  filter(adi_decile == 1 | adi_decile == 10) |> 
  group_by(PX_ID) |> 
  tbl_summary(by = adi_decile, label = list(
    can_race ~ "Race",
    CAN_GENDER ~ "Sex"
  ),
  statistic = list(all_continuous() ~ "{mean} ({sd})", all_categorical() ~ "{n} ({p}%)")) |> 
  modify_table_body(filter, !(variable == "PX_ID" ))

heart_SVM |> 
  select(CAN_AGE_AT_LISTING, CAN_GENDER, can_race, durable_LVAD, temporary_LVAD, RVAD, LVAD_RVAD, TAH, initial_status, SVM_PERCENTILE, SVI_percentile, mean_ADI, dci_percentile, status, CAN_ETHNICITY_SRTR ) |> 
  group_by(PX_ID) |> 
  tbl_summary(label = list(
    can_race ~ "Race",
    CAN_GENDER ~ "Sex"
  ),
  statistic = list(all_continuous() ~ "{mean} ({sd})", all_categorical() ~ "{n} ({p}%)"))
```

## Regression Models

SVM Transplant

```{r SVM crr transplant}
heart_SVM <- heart_SVM |> 
  mutate(sex = case_when(
    CAN_GENDER == "M" ~ 1,
    CAN_GENDER == "F" ~ 0),
    hispanic = case_when(
      CAN_ETHNICITY_SRTR == "LATINO" ~ 1,
      CAN_ETHNICITY_SRTR == "NLATIN" ~ 0,
      .default = NA),
    initial_status_numeric = case_when(
      CAN_INIT_STAT %in% c('1010', '1020', '1090', '1110', '2010', '2020', '2090', '2110', '3010', '6010', '6011', '6012', '9010') ~ 1,
           CAN_INIT_STAT %in% c('1030', '1120', '2030', '2120', '6002', '6020', '6030', '9020') ~ 2,
           CAN_INIT_STAT %in% c('1130', '2130', '6040') ~ 3,
           CAN_INIT_STAT %in% c('1140', '2140', '6004') ~ 4,
           CAN_INIT_STAT %in% c('1150', '2150') ~ 5,
           CAN_INIT_STAT %in% c('1160', '2160') ~ 6,
           .default = NA),
    status_numeric = case_when(
      status == "censor" ~ 0,
      status == "transplant" ~ 1,
      status == "death/deteriorate" ~ 2
    ))
```

## SVM CRR stratified by center 

```{r svm crrs prep, echo = TRUE, results = 'hide' }
cov.test <- model.matrix(~ factor(heart_SVM$svm_decile) +
                  heart_SVM$durable_LVAD +
                  factor(heart_SVM$can_race) +
                  heart_SVM$sex +
                  heart_SVM$hispanic +
                  heart_SVM$CAN_AGE_AT_LISTING +
                  heart_SVM$temporary_LVAD +
                  heart_SVM$RVAD +
                  heart_SVM$TAH +
                  heart_SVM$LVAD_RVAD + 
                  factor(heart_SVM$initial_status_numeric))[,-1]
cov.test <- cov.test[complete.cases(cov.test), ]
complete_cases <- heart_SVM[complete.cases(heart_SVM[, c("svm_decile", "durable_LVAD", "can_race", "sex", "hispanic", "CAN_AGE_AT_LISTING", "temporary_LVAD", "RVAD", "TAH", "LVAD_RVAD", "initial_status_numeric")]), ]
```

1-9 = 2-10 decile
10= durable lvad
11 - 15 = race (other, black, asian, native american, native Hawaiian)
16 - sex
17 - hispanic
18 - age
19 - temporary lvad
20 - rvad
21 - tah
22 - lvad rvad
23-27 = status (2-6)


```{r svm crrs}
svm_crrs <- crrs(ftime = complete_cases$survival_time, fstatus = complete_cases$status_numeric, cov1 = cov.test, strata = complete_cases$CAN_LISTING_CTR_CD, ctype = 1)
svm_crrs
```

## SVI CRR stratified by center Z

```{r svi crrs prep, echo = TRUE, results = 'hide' }
cov.test <- model.matrix(~ factor(heart_SVM$svi_decile) +
                  heart_SVM$durable_LVAD +
                  factor(heart_SVM$can_race) +
                  heart_SVM$sex +
                  heart_SVM$hispanic +
                  heart_SVM$CAN_AGE_AT_LISTING +
                  heart_SVM$temporary_LVAD +
                  heart_SVM$RVAD +
                  heart_SVM$TAH +
                  heart_SVM$LVAD_RVAD + 
                  factor(heart_SVM$initial_status_numeric))[,-1]
cov.test <- cov.test[complete.cases(cov.test), ]
complete_cases <- heart_SVM[complete.cases(heart_SVM[, c("svi_decile", "durable_LVAD", "can_race", "sex", "hispanic", "CAN_AGE_AT_LISTING", "temporary_LVAD", "RVAD", "TAH", "LVAD_RVAD", "initial_status_numeric")]), ]
```

```{r svi crrs}
svi_crrs <- crrs(ftime = complete_cases$survival_time, fstatus = complete_cases$status_numeric, cov1 = cov.test, strata = complete_cases$CAN_LISTING_CTR_CD, ctype = 1)
svi_crrs

```


## ADI CRR stratified by center Z

```{r adi crrs prep, echo = TRUE, results = 'hide' }
cov.test <- model.matrix(~ factor(heart_SVM$adi_decile) +
                  heart_SVM$durable_LVAD +
                  factor(heart_SVM$can_race) +
                  heart_SVM$sex +
                  heart_SVM$hispanic +
                  heart_SVM$CAN_AGE_AT_LISTING +
                  heart_SVM$temporary_LVAD +
                  heart_SVM$RVAD +
                  heart_SVM$TAH +
                  heart_SVM$LVAD_RVAD + 
                  factor(heart_SVM$initial_status_numeric))[,-1]
cov.test <- cov.test[complete.cases(cov.test), ]
complete_cases <- heart_SVM[complete.cases(heart_SVM[, c("adi_decile", "durable_LVAD", "can_race", "sex", "hispanic", "CAN_AGE_AT_LISTING", "temporary_LVAD", "RVAD", "TAH", "LVAD_RVAD", "initial_status_numeric")]), ]
```

```{r adi crrs}
adi_crrs <- crrs(ftime = complete_cases$survival_time, fstatus = complete_cases$status_numeric, cov1 = cov.test, strata = complete_cases$CAN_LISTING_CTR_CD, ctype = 1)
adi_crrs
```


## DCI CRR stratified by center Z

```{r dci crrs prep, echo = TRUE, results = 'hide' }
cov.test <- model.matrix(~ factor(heart_SVM$dci_decile) +
                  heart_SVM$durable_LVAD +
                  factor(heart_SVM$can_race) +
                  heart_SVM$sex +
                  heart_SVM$hispanic +
                  heart_SVM$CAN_AGE_AT_LISTING +
                  heart_SVM$temporary_LVAD +
                  heart_SVM$RVAD +
                  heart_SVM$TAH +
                  heart_SVM$LVAD_RVAD + 
                  factor(heart_SVM$initial_status_numeric))[,-1]
cov.test <- cov.test[complete.cases(cov.test), ]
complete_cases <- heart_SVM[complete.cases(heart_SVM[, c("dci_decile", "durable_LVAD", "can_race", "sex", "hispanic", "CAN_AGE_AT_LISTING", "temporary_LVAD", "RVAD", "TAH", "LVAD_RVAD", "initial_status_numeric")]), ]
```


```{r dci crrs}
dci_crrs <- crrs(ftime = complete_cases$survival_time, fstatus = complete_cases$status_numeric, cov1 = cov.test, strata = complete_cases$CAN_LISTING_CTR_CD, ctype = 1)
dci_crrs
```
 


```{r svm tidy crr}
library(tidycmprsk)

svm <- crr(Surv(survival_time,status) ~ SVM_decile + can_race + ethnicity +  CAN_GENDER + CAN_AGE_AT_LISTING + durable_LVAD +temporary_LVAD + RVAD + TAH + LVAD_RVAD + initial_status, data = heart_SVM)


svm_death <- crr(Surv(survival_time,status) ~ SVM_decile + can_race + ethnicity +  CAN_GENDER + CAN_AGE_AT_LISTING + durable_LVAD +temporary_LVAD + RVAD + TAH + LVAD_RVAD + initial_status, data = heart_SVM, failcode = "death/deteriorate")

tbl1<- svm |> 
  tbl_regression(exponentiate = TRUE, label = list(
    SVM_decile ~ "SVM Decile",
    can_race ~ "Race",
    ethnicity ~ "Ethnicity",
    CAN_GENDER ~ "Sex",
    CAN_AGE_AT_LISTING ~ "Age",
    durable_LVAD ~ "Durable LVAD",
    temporary_LVAD ~ "Temporary LVAD",
    LVAD_RVAD ~ "LVAD + RVAD",
    initial_status ~ "Initial Status"
  )) |> 
  modify_header(label = "") 
tbl2 <- svm_death |> tbl_regression(exponentiate = TRUE, label = list(
    SVM_decile ~ "SVM Decile",
    can_race ~ "Race",
    ethnicity ~ "Ethnicity",
    CAN_GENDER ~ "Sex",
    CAN_AGE_AT_LISTING ~ "Age",
    durable_LVAD ~ "Durable LVAD",
    temporary_LVAD ~ "Temporary LVAD",
    LVAD_RVAD ~ "LVAD + RVAD",
    initial_status ~ "Initial Status"
  )) |> 
  modify_header(label = "") 

  tbl_merge(
    tbls = list(tbl1, tbl2),
    tab_spanner = c("**Time to Transplant**", "**Time to Death or Deterioration**")
  )


```


SVI Transplant

```{r SVI crr transplant}

svi <- crr(Surv(survival_time,status) ~ SVI_decile + can_race + ethnicity +  CAN_GENDER + CAN_AGE_AT_LISTING + durable_LVAD +temporary_LVAD + RVAD + TAH + LVAD_RVAD + initial_status, data = heart_SVM)



svi_death <- crr(Surv(survival_time,status) ~ SVI_decile + can_race + ethnicity +  CAN_GENDER + CAN_AGE_AT_LISTING + durable_LVAD +temporary_LVAD + RVAD + TAH + LVAD_RVAD + initial_status, data = heart_SVM, failcode = "death/deteriorate")

tbl1<- svi |> 
  tbl_regression(exponentiate = TRUE, label = list(
    SVI_decile ~ "SVI Decile",
    can_race ~ "Race",
    ethnicity ~ "Ethnicity",
    CAN_GENDER ~ "Sex",
    CAN_AGE_AT_LISTING ~ "Age",
    durable_LVAD ~ "Durable LVAD",
    temporary_LVAD ~ "Temporary LVAD",
    LVAD_RVAD ~ "LVAD + RVAD",
    initial_status ~ "Initial Status"
  )) |> 
  modify_header(label = "") 
tbl2 <- svi_death |> tbl_regression(exponentiate = TRUE, label = list(
    SVI_decile ~ "SVI Decile",
    can_race ~ "Race",
    ethnicity ~ "Ethnicity",
    CAN_GENDER ~ "Sex",
    CAN_AGE_AT_LISTING ~ "Age",
    durable_LVAD ~ "Durable LVAD",
    temporary_LVAD ~ "Temporary LVAD",
    LVAD_RVAD ~ "LVAD + RVAD",
    initial_status ~ "Initial Status"
  )) |> 
  modify_header(label = "") 

  tbl_merge(
    tbls = list(tbl1, tbl2),
    tab_spanner = c("**Time to Transplant**", "**Time to Death or Deterioration**")
  )




```

ADI Transplant

```{r crr ADI transplant}


adi <- crr(Surv(survival_time,status) ~ ADI_decile + can_race + ethnicity +  CAN_GENDER + CAN_AGE_AT_LISTING + durable_LVAD +temporary_LVAD + RVAD + TAH + LVAD_RVAD + initial_status, data = heart_SVM)


adi_death <- crr(Surv(survival_time,status) ~ ADI_decile + can_race + ethnicity +  CAN_GENDER + CAN_AGE_AT_LISTING + durable_LVAD +temporary_LVAD + RVAD + TAH + LVAD_RVAD + initial_status, data = heart_SVM, failcode = "death/deteriorate")

tbl1<- adi |> 
  tbl_regression(exponentiate = TRUE, label = list(
    ADI_decile ~ "ADI Decile",
    can_race ~ "Race",
    ethnicity ~ "Ethnicity",
    CAN_GENDER ~ "Sex",
    CAN_AGE_AT_LISTING ~ "Age",
    durable_LVAD ~ "Durable LVAD",
    temporary_LVAD ~ "Temporary LVAD",
    LVAD_RVAD ~ "LVAD + RVAD",
    initial_status ~ "Initial Status"
  )) |> 
  modify_header(label = "") 
tbl2 <- adi_death |> tbl_regression(exponentiate = TRUE, label = list(
    ADI_decile ~ "ADI Decile",
    can_race ~ "Race",
    ethnicity ~ "Ethnicity",
    CAN_GENDER ~ "Sex",
    CAN_AGE_AT_LISTING ~ "Age",
    durable_LVAD ~ "Durable LVAD",
    temporary_LVAD ~ "Temporary LVAD",
    LVAD_RVAD ~ "LVAD + RVAD",
    initial_status ~ "Initial Status"
  )) |> 
  modify_header(label = "") 

  tbl_merge(
    tbls = list(tbl1, tbl2),
    tab_spanner = c("**Time to Transplant**", "**Time to Death or Deterioration**")
  )

```

DCI Transplant 

```{r dci tidy crr}

dci <- crr(Surv(survival_time,status) ~ DCI_decile + can_race + ethnicity +  CAN_GENDER + CAN_AGE_AT_LISTING + durable_LVAD +temporary_LVAD + RVAD + TAH + LVAD_RVAD + initial_status, data = heart_SVM)


dci_death <- crr(Surv(survival_time,status) ~ DCI_decile + can_race + ethnicity +  CAN_GENDER + CAN_AGE_AT_LISTING + durable_LVAD +temporary_LVAD + RVAD + TAH + LVAD_RVAD + initial_status, data = heart_SVM, failcode = "death/deteriorate")


tbl1<- dci |> 
  tbl_regression(exponentiate = TRUE, label = list(
    DCI_decile ~ "DCI Decile",
    can_race ~ "Race",
    ethnicity ~ "Ethnicity",
    CAN_GENDER ~ "Sex",
    CAN_AGE_AT_LISTING ~ "Age",
    durable_LVAD ~ "Durable LVAD",
    temporary_LVAD ~ "Temporary LVAD",
    LVAD_RVAD ~ "LVAD + RVAD",
    initial_status ~ "Initial Status"
  )) |> 
  modify_header(label = "") 
tbl2 <- dci_death |> tbl_regression(exponentiate = TRUE, label = list(
    DCI_decile ~ "DCI Decile",
    can_race ~ "Race",
    ethnicity ~ "Ethnicity",
    CAN_GENDER ~ "Sex",
    CAN_AGE_AT_LISTING ~ "Age",
    durable_LVAD ~ "Durable LVAD",
    temporary_LVAD ~ "Temporary LVAD",
    LVAD_RVAD ~ "LVAD + RVAD",
    initial_status ~ "Initial Status"
  )) |> 
  modify_header(label = "") 

  tbl_merge(
    tbls = list(tbl1, tbl2),
    tab_spanner = c("**Time to Transplant**", "**Time to Death or Deterioration**")
  )

```


## Forest Plots

```{r svm forest}


forest <- tidy(svm)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)

svm_forest <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  xlim(0.6,1.2)+
  labs(x = "Sub-Hazard Ratio of Transplantation",
       y = "SVM Decile")+
  theme_classic()


forest <- tidy(svm_death)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)

svm_death_forest <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  labs(x = "Sub-Hazard Ratio of Death or Deterioration",
       y = "SVM Decile")+
   xlim(0.25, 3.5)+
  theme_classic()
```

```{r svi forest}

forest <- tidy(svi)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)

svi_forest <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  xlim(0.6,1.2)+
  labs(x = "Sub-Hazard Ratio of Transplantation",
       y = "SVI Decile")+
  theme_classic()

forest <- tidy(svi_death)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)

svi_death_forest <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  labs(x = "Sub-Hazard Ratio of Death or Deterioration",
       y = "SVI Decile")+
   xlim(0.25, 3.5)+
  theme_classic()
```


```{r adi forest}

forest <- tidy(adi)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)

adi_forest <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point(color = "black") +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
    xlim(0.6,1.2)+
  labs(x = "Sub-Hazard Ratio of Transplantation",
       y = "ADI Decile")+
  theme_classic()

forest <- tidy(adi_death)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)

adi_death_forest <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point(color = "black") +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  labs(x = "Sub-Hazard Ratio of Death or Deterioration",
       y = "ADI Decile")+
  xlim(0.25, 3.5)+
  theme_classic()
```


```{r dci forest}

forest <- tidy(dci)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)

dci_forest <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
    xlim(0.6,1.2)+
  labs(x = "Sub-Hazard Ratio of Transplantation",
       y = "DCI Decile")+
  theme_classic()

forest <- tidy(dci_death)
print(forest)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))
forest <- forest[c(1:9),]
forest <- forest |>
  mutate(decile = as.numeric(sub(".*?([0-9]+)[a-zA-Z ]*", "\\1",term))) |> 
  arrange(decile)
forest$term <- factor(forest$term, levels = forest$term)
forest$decile <- as.factor(forest$decile)

dci_death_forest <- forest |> 
   ggplot(aes(x = hazard, y = decile)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  labs(x = "Sub-Hazard Ratio of Death or Deterioration",
       y = "DCI Decile")+
  xlim(0.25, 3.5)+
  theme_classic()
```

```{r figure 3}
grid.arrange(adi_forest,adi_death_forest, ncol =2, nrow = 1)

grid.arrange(svm_forest,svm_death_forest, svi_forest,svi_death_forest, adi_forest,adi_death_forest, dci_forest,dci_death_forest, ncol = 2, nrow = 4)

```


## Spline Plots

Fail code for cumulative indidence is transplant. 2 cohorts: durable LVAD or no durable lvad

For prediction, set equal to the median age

Set ADI to be the case

objective of this is to equalize transplant across transplant. The absolute number of place based points, type of rating scale, shape of the rating scale will need to be determined through simulation optimization

Summer Gentry paper

```{r coxme cindex}

required_vars <- c("survival_time", "transplant_outcome", "SVM_PERCENTILE", "CAN_LISTING_CTR_CD")
complete_cases <- complete.cases(heart_SVM[, required_vars])
heart_SVM_complete <- heart_SVM[complete_cases, ]
SVM_model_transplant <- coxme(Surv(survival_time,transplant_outcome) ~ rcs(SVM_PERCENTILE) + (1|CAN_LISTING_CTR_CD), data = heart_SVM_complete)
linear_predictor <- -1 * SVM_model_transplant$linear.predictor
concordance(Surv(heart_SVM_complete$survival_time, heart_SVM_complete$transplant_outcome) ~ linear_predictor)


required_vars <- c("survival_time", "transplant_outcome", "SVI_percentile", "CAN_LISTING_CTR_CD")
complete_cases <- complete.cases(heart_SVM[, required_vars])
heart_SVM_complete <- heart_SVM[complete_cases, ]
SVI_model_transplant <- coxme(Surv(survival_time,transplant_outcome) ~ rcs(SVI_percentile) + (1|CAN_LISTING_CTR_CD), data = heart_SVM_complete)
linear_predictor <- -1 * SVI_model_transplant$linear.predictor
concordance(Surv(heart_SVM_complete$survival_time, heart_SVM_complete$transplant_outcome) ~ linear_predictor)

required_vars <- c("survival_time", "transplant_outcome", "mean_ADI", "CAN_LISTING_CTR_CD")
complete_cases <- complete.cases(heart_SVM[, required_vars])
heart_SVM_complete <- heart_SVM[complete_cases, ]
ADI_model_transplant <- coxme(Surv(survival_time,transplant_outcome) ~ rcs(mean_ADI) + (1|CAN_LISTING_CTR_CD), data = heart_SVM_complete)
linear_predictor <- -1 * ADI_model_transplant$linear.predictor
concordance(Surv(heart_SVM_complete$survival_time, heart_SVM_complete$transplant_outcome) ~ linear_predictor)

required_vars <- c("survival_time", "transplant_outcome", "dci_percentile", "CAN_LISTING_CTR_CD")
complete_cases <- complete.cases(heart_SVM[, required_vars])
heart_SVM_complete <- heart_SVM[complete_cases, ]
DCI_model_transplant <- coxme(Surv(survival_time,transplant_outcome) ~ rcs(dci_percentile) + (1|CAN_LISTING_CTR_CD), data = heart_SVM_complete)
linear_predictor <- -1 * DCI_model_transplant$linear.predictor
concordance(Surv(heart_SVM_complete$survival_time, heart_SVM_complete$transplant_outcome) ~ linear_predictor)

```


```{r AUC}
set.seed(1234)
SVM_model_transplant <- coxme(Surv(survival_time,transplant_outcome) ~ rcs(SVM_PERCENTILE) + (1|CAN_LISTING_CTR_CD), data = heart_SVM)
SVI_model_transplant<- coxme(Surv(survival_time,transplant_outcome) ~ rcs(SVI_percentile) + (1|CAN_LISTING_CTR_CD), data = heart_SVM)
ADI_model_transplant <- coxme(Surv(survival_time,transplant_outcome) ~ rcs(mean_ADI)+ (1|CAN_LISTING_CTR_CD), data = heart_SVM)
DCI_model_transplant <- coxme(Surv(survival_time,transplant_outcome) ~ rcs(dci_percentile)+ (1|CAN_LISTING_CTR_CD), data = heart_SVM)

SVM_model_transplant
SVI_model_transplant
ADI_model_transplant
DCI_model_transplant

zci <- function(fit, p=.95) {
ilogist <- function(p) log(p/(1-p)) # inverse logistic
logistic <- function(x) exp(x)/(1 + exp(x))
temp <- concordance(fit, influence =1)
cminus <- temp$concordance - temp$dfbeta # values of concordance, without i
newd <- ilogist(temp$concordance) - ilogist(cminus) # dfbeta on new scale
new.sd <- sqrt(sum(newd^2))
z <- qnorm((1-p)/2)
new.ci <- logistic(ilogist(temp$concordance) + c(z, -z)* new.sd)
rbind( new= new.ci)
}

required_vars <- c("survival_time", "transplant_outcome", "SVM_PERCENTILE", "CAN_LISTING_CTR_CD")
complete_cases <- complete.cases(heart_SVM[, required_vars])
heart_SVM_complete <- heart_SVM[complete_cases, ]
SVM_model_transplant <- coxme(Surv(survival_time,transplant_outcome) ~ rcs(SVM_PERCENTILE) + (1|CAN_LISTING_CTR_CD), data = heart_SVM_complete)
linear_predictor <- SVM_model_transplant$linear.predictor
concordance(Surv(heart_SVM_complete$survival_time, heart_SVM_complete$transplant_outcome) ~ linear_predictor)

```


```{r transplant spline}
SVM_model_transplant <- coxph(Surv(survival_time,transplant_outcome) ~ rcs(SVM_PERCENTILE), data = heart_SVM)
SVI_model_transplant<- coxph(Surv(survival_time,transplant_outcome) ~ rcs(SVI_percentile), data = heart_SVM)
ADI_model_transplant <- coxph(Surv(survival_time,transplant_outcome) ~ rcs(mean_ADI), data = heart_SVM)
DCI_model_transplant <- coxph(Surv(survival_time,transplant_outcome) ~ rcs(dci_percentile), data = heart_SVM)

avg_patient <- tibble(
  SVM_PERCENTILE = rep(seq(0,100),2))
avg_patient1 <- tibble(
  SVI_percentile = rep(seq(0,100),2))
avg_patient2 <- tibble(
  mean_ADI = rep(seq(0,100),2))
avg_patient3 <- tibble(
  dci_percentile = rep(seq(0,100),2))

hazard_transplant_SVM <- predict(SVM_model_transplant, newdata = avg_patient)
hazard_transplant_SVI <- predict(SVI_model_transplant, newdata = avg_patient1)
hazard_transplant_ADI <- predict(ADI_model_transplant, newdata = avg_patient2)
hazard_transplant_DCI <- predict(DCI_model_transplant, newdata = avg_patient3)

change_basecase_to_first_percentile <- function(hazard_list){
  log_haz_1st_SVM <- hazard_list[[50]]

hazard_list<- hazard_list - log_haz_1st_SVM

exp(hazard_list)
  
}
hazard_transplant_SVM <- change_basecase_to_first_percentile(hazard_transplant_SVM)
hazard_transplant_SVI <- change_basecase_to_first_percentile(hazard_transplant_SVI)
hazard_transplant_ADI <- change_basecase_to_first_percentile(hazard_transplant_ADI)
hazard_transplant_DCI <- change_basecase_to_first_percentile(hazard_transplant_DCI)


spline_transplant <- avg_patient |> 
  cbind(hazard_transplant_SVM,
        hazard_transplant_SVI,
        hazard_transplant_ADI, 
        hazard_transplant_DCI) |> 
  ggplot()+
  geom_smooth(aes(x=SVM_PERCENTILE, y = hazard_transplant_SVM, color = "SVM"), data = avg_patient) +
    geom_smooth(aes(x=SVI_percentile, y = hazard_transplant_SVI, color = "SVI"), data = avg_patient1)+
    geom_smooth(aes(x=mean_ADI, y = hazard_transplant_ADI, color = "ADI"), data = avg_patient2)+
    geom_smooth(aes(x=dci_percentile, y = hazard_transplant_DCI, color = "DCI"), data = avg_patient3)+
  xlab("SDoH Composite Score Percentile")+
  ylab("Cause Specific HR of Transplant")+
  theme(legend.title = element_blank(),
        legend.position = "bottom")

spline_transplant
  
```


```{r death spline}
SVM_model_death <- coxph(Surv(survival_time,death_outcome) ~ rcs(SVM_PERCENTILE), data = heart_SVM)
SVI_model_death <- coxph(Surv(survival_time,death_outcome) ~ rcs(SVI_percentile) , data = heart_SVM)
ADI_model_death <- coxph(Surv(survival_time,death_outcome) ~ rcs(mean_ADI), data = heart_SVM)
DCI_model_death <- coxph(Surv(survival_time,death_outcome) ~ rcs(dci_percentile), data = heart_SVM)


avg_patient <- tibble(
  SVM_PERCENTILE = rep(seq(0,100),2))
avg_patient1 <- tibble(
  SVI_percentile = rep(seq(0,100),2))
avg_patient2 <- tibble(
  mean_ADI = rep(seq(0,100),2))
avg_patient3 <- tibble(
  dci_percentile = rep(seq(0,100),2))


hazard_death_SVM <- predict(SVM_model_death, newdata = avg_patient)
hazard_death_SVI <- predict(SVI_model_death, newdata = avg_patient1)
hazard_death_ADI <- predict(ADI_model_death, newdata = avg_patient2)
hazard_death_DCI <- predict(DCI_model_death, newdata = avg_patient3)


change_basecase_to_first_percentile <- function(hazard_list){
  log_haz_1st_SVM <- hazard_list[[50]]

hazard_list<- hazard_list - log_haz_1st_SVM

exp(hazard_list)
  
}
hazard_death_SVM <- change_basecase_to_first_percentile(hazard_death_SVM)
hazard_death_SVI <- change_basecase_to_first_percentile(hazard_death_SVI)
hazard_death_ADI <- change_basecase_to_first_percentile(hazard_death_ADI)
hazard_death_DCI <- change_basecase_to_first_percentile(hazard_death_DCI)



spline_death <- avg_patient |> 
  cbind(hazard_death_SVM,
        hazard_death_SVI,
        hazard_death_ADI, 
        hazard_death_DCI) |> 
  ggplot()+
  geom_smooth(aes(x=SVM_PERCENTILE, y = hazard_death_SVM, color = "SVM"), data = avg_patient) +
    geom_smooth(aes(x=SVI_percentile, y = hazard_death_SVI, color = "SVI"), data = avg_patient1)+
    geom_smooth(aes(x=mean_ADI, y = hazard_death_ADI, color = "ADI"), data = avg_patient2)+
    geom_smooth(aes(x=dci_percentile, y = hazard_death_DCI, color = "DCI"), data = avg_patient3)+
  xlab("SDoH Composite Score Percentile")+
  ylab("Cause Specific HR of Death or Deterioration")+
  theme(legend.title = element_blank(),
        legend.position = "bottom")

spline_death
  
```



```{r odds of LVAD}
SVM_model_LVAD <- glm(durable_LVAD ~ rcs(SVM_PERCENTILE), data = heart_SVM, family = "binomial")
SVI_model_LVAD <- glm(durable_LVAD ~ rcs(SVI_percentile), data = heart_SVM, family = "binomial")
ADI_model_LVAD <- glm(durable_LVAD ~ rcs(mean_ADI), data = heart_SVM, family = "binomial")
DCI_model_LVAD <- glm(durable_LVAD ~ rcs(dci_percentile), data = heart_SVM, family = "binomial")

concordance(SVM_model_LVAD)
zci(SVM_model_LVAD)

concordance(SVI_model_LVAD)
zci(SVI_model_LVAD)


concordance(ADI_model_LVAD)
zci(ADI_model_LVAD)


concordance(DCI_model_LVAD)
zci(DCI_model_LVAD)

avg_patient <- tibble(
  SVM_PERCENTILE = rep(seq(0,100),2),
  durable_LVAD = c(rep(0,101), rep(1,101)))
avg_patient1 <- tibble(
  SVI_percentile = rep(seq(0,100),2),
  durable_LVAD = c(rep(0,101), rep(1,101)))
avg_patient2 <- tibble(
  mean_ADI = rep(seq(0,100),2),
  durable_LVAD = c(rep(0,101), rep(1,101)))
avg_patient3 <- tibble(
  dci_percentile = rep(seq(0,100),2),
  durable_LVAD = c(rep(0,101), rep(1,101)))


odds_LVAD_SVM <- predict(SVM_model_LVAD, newdata = avg_patient)
odds_LVAD_SVI <- predict(SVI_model_LVAD, newdata = avg_patient1)
odds_LVAD_ADI <- predict(ADI_model_LVAD, newdata = avg_patient2)
odds_LVAD_DCI <- predict(DCI_model_LVAD, newdata = avg_patient3)


change_basecase_to_first_percentile <- function(odds_list){
  log_haz_1st_SVM <- odds_list[[50]]

odds_list<- odds_list - log_haz_1st_SVM

exp(odds_list)
  
}
odds_LVAD_SVM <- change_basecase_to_first_percentile(odds_LVAD_SVM)
odds_LVAD_SVI <- change_basecase_to_first_percentile(odds_LVAD_SVI)
odds_LVAD_ADI <- change_basecase_to_first_percentile(odds_LVAD_ADI)
odds_LVAD_DCI <- change_basecase_to_first_percentile(odds_LVAD_DCI)



spline_lvad <- avg_patient |> 
  cbind(odds_LVAD_SVM,
        odds_LVAD_SVI,
        odds_LVAD_ADI, 
        odds_LVAD_DCI) |> 
  ggplot()+
  geom_smooth(aes(x=SVM_PERCENTILE, y = odds_LVAD_SVM, color = "SVM", ), data = avg_patient) +
    geom_smooth(aes(x=SVI_percentile, y = odds_LVAD_SVI, color = "SVI"), data = avg_patient1)+
    geom_smooth(aes(x=mean_ADI, y = odds_LVAD_ADI, color = "ADI"), data = avg_patient2)+
    geom_smooth(aes(x=dci_percentile, y = odds_LVAD_DCI, color = "DCI"), data = avg_patient3)+
  xlab("SDoH Composite Score Percentile")+
  ylab("Odds Ratio of Durable LVAD")+
  theme(legend.title = element_blank(),
        legend.position = "bottom")

  spline_lvad
```

```{r}
grid.arrange(spline_transplant, spline_death, spline_lvad, ncol=3 )
grid.arrange(spline_transplant, spline_death, ncol=2 )

```



```{r cumic curves}
heart_SVM$statuscuminc <- factor(heart_SVM$status_numeric,
                           levels = c(0,1,2),
                           labels = c("On waitlist at end of follow-up","Received transplant","Died or removed for deterioration"))

svm_cuminc <- cuminc(Surv(survival_time,statuscuminc) ~ SVM_decile, data = subset(heart_SVM, (SVM_decile == "1st Decile" | SVM_decile == "10th Decile"))) %>% 
  ggcuminc(outcome = c("Received transplant","Died or removed for deterioration")) +
  ylim(0,1)+
  labs(
    x = "Days",
    title = "SVM") + 
  add_confidence_interval()+
  theme(plot.title = element_text(hjust=0.5))


svi_cuminc <- cuminc(Surv(survival_time,statuscuminc) ~ SVI_decile, data = subset(heart_SVM, (SVI_decile == "1st Decile" | SVI_decile == "10th Decile"))) %>% 
  ggcuminc(outcome = c("Received transplant","Died or removed for deterioration")) +
    ylim(0,1)+
  labs(
    x = "Days",
    title = "SVI") + 
  add_confidence_interval()+
  theme(plot.title = element_text(hjust=0.5))


adi_cuminc <- cuminc(Surv(survival_time,statuscuminc) ~ ADI_decile, data = subset(heart_SVM, (ADI_decile == "1st Decile" | ADI_decile == "10th Decile"))) %>% 
  ggcuminc(outcome = c("Received transplant","Died or removed for deterioration")) +
    ylim(0,1)+
labs(
    x = "Days",
    title = "ADI") + 
  add_confidence_interval()+
  theme(plot.title = element_text(hjust=0.5))


dci_cuminc <- cuminc(Surv(survival_time,statuscuminc) ~ DCI_decile, data = subset(heart_SVM, (DCI_decile == "1st Decile" | DCI_decile == "10th Decile"))) %>% 
  ggcuminc(outcome = c("Received transplant","Died or removed for deterioration")) +
   ylim(0,1)+
labs(
    x = "Days",
    title = "DCI") + 
  add_confidence_interval()+
  theme(plot.title = element_text(hjust=0.5))

grid.arrange(svm_cuminc, svi_cuminc, adi_cuminc, dci_cuminc, ncol=2, nrow = 2 )
```

```{r}
set.seed(1234)
SVM_model_transplant <- coxme(Surv(survival_time,transplant_outcome) ~ rcs(SVM_PERCENTILE) + (1|CAN_LISTING_CTR_CD), data = heart_SVM)
SVI_model_transplant<- coxme(Surv(survival_time,transplant_outcome) ~ rcs(SVI_percentile) + (1|CAN_LISTING_CTR_CD), data = heart_SVM)
ADI_model_transplant <- coxme(Surv(survival_time,transplant_outcome) ~ rcs(mean_ADI)+ (1|CAN_LISTING_CTR_CD), data = heart_SVM)
DCI_model_transplant <- coxme(Surv(survival_time,transplant_outcome) ~ rcs(dci_percentile)+ (1|CAN_LISTING_CTR_CD), data = heart_SVM)

avg_patient <- tibble(
  SVM_PERCENTILE = rep(seq(0,100)))
avg_patient1 <- tibble(
  SVI_percentile = rep(seq(0,100)))
avg_patient2 <- tibble(
  mean_ADI = rep(seq(0,100)))
avg_patient3 <- tibble(
  dci_percentile = rep(seq(0,100)))


hazard_transplant_SVM <- predict_coxme(SVM_model_transplant, newdata = avg_patient, type = c("risk"))
hazard_transplant_SVI <- predict_coxme(SVI_model_transplant, newdata = avg_patient1, type = c("risk"))
hazard_transplant_ADI <- predict_coxme(ADI_model_transplant, newdata = avg_patient2, type = c("risk"))
hazard_transplant_DCI <- predict_coxme(DCI_model_transplant, newdata = avg_patient3, type = c("risk"))



change_basecase_to_first_percentile <- function(hazard_list){
  log_haz_1st_SVM <- hazard_list[[52]]

hazard_list<- hazard_list - log_haz_1st_SVM

exp(hazard_list)
  
}
hazard_transplant_SVM <- change_basecase_to_first_percentile(hazard_transplant_SVM)



hazard_transplant_SVM <- change_basecase_to_first_percentile(hazard_transplant_SVM)
hazard_transplant_SVI <- change_basecase_to_first_percentile(hazard_transplant_SVI)
hazard_transplant_ADI <- change_basecase_to_first_percentile(hazard_transplant_ADI)
hazard_transplant_DCI <- change_basecase_to_first_percentile(hazard_transplant_DCI)


spline_transplant <- avg_patient |> 
  cbind(hazard_transplant_SVM,
        hazard_transplant_SVI,
        hazard_transplant_ADI, 
        hazard_transplant_DCI) |> 
  ggplot()+
  geom_smooth(aes(x=SVM_PERCENTILE, y = hazard_transplant_SVM, color = "SVM"), data = avg_patient) +
    geom_smooth(aes(x=SVI_percentile, y = hazard_transplant_SVI, color = "SVI"), data = avg_patient1)+
    geom_smooth(aes(x=mean_ADI, y = hazard_transplant_ADI, color = "ADI"), data = avg_patient2)+
    geom_smooth(aes(x=dci_percentile, y = hazard_transplant_DCI, color = "DCI"), data = avg_patient3)+
  xlab("SDoH Composite Score Percentile")+
  ylab("Cause Specific HR of Transplant")+
  theme(legend.title = element_blank(),
        legend.position = "bottom")

spline_transplant

```

```{r}
set.seed(1234)
SVM_model_death <- coxme(Surv(survival_time,death_outcome) ~ rcs(SVM_PERCENTILE) + (1|CAN_LISTING_CTR_CD), data = heart_SVM)
SVI_model_death<- coxme(Surv(survival_time,death_outcome) ~ rcs(SVI_percentile) + (1|CAN_LISTING_CTR_CD), data = heart_SVM)
ADI_model_death <- coxme(Surv(survival_time,death_outcome) ~ rcs(mean_ADI)+ (1|CAN_LISTING_CTR_CD), data = heart_SVM)
DCI_model_death <- coxme(Surv(survival_time,death_outcome) ~ rcs(dci_percentile)+ (1|CAN_LISTING_CTR_CD), data = heart_SVM)

avg_patient <- tibble(
  SVM_PERCENTILE = rep(seq(0,100)))
avg_patient1 <- tibble(
  SVI_percentile = rep(seq(0,100)))
avg_patient2 <- tibble(
  mean_ADI = rep(seq(0,100)))
avg_patient3 <- tibble(
  dci_percentile = rep(seq(0,100)))


hazard_death_SVM <- predict_coxme(SVM_model_death, newdata = avg_patient, type = c("risk"))
hazard_death_SVI <- predict_coxme(SVI_model_death, newdata = avg_patient1, type = c("risk"))
hazard_death_ADI <- predict_coxme(ADI_model_death, newdata = avg_patient2, type = c("risk"))
hazard_death_DCI <- predict_coxme(DCI_model_death, newdata = avg_patient3, type = c("risk"))



change_basecase_to_first_percentile <- function(hazard_list){
  log_haz_1st_SVM <- hazard_list[[52]]

hazard_list<- hazard_list - log_haz_1st_SVM

exp(hazard_list)
  
}
hazard_death_SVM <- change_basecase_to_first_percentile(hazard_death_SVM)



hazard_death_SVM <- change_basecase_to_first_percentile(hazard_death_SVM)
hazard_death_SVI <- change_basecase_to_first_percentile(hazard_death_SVI)
hazard_death_ADI <- change_basecase_to_first_percentile(hazard_death_ADI)
hazard_death_DCI <- change_basecase_to_first_percentile(hazard_death_DCI)


spline_death <- avg_patient |> 
  cbind(hazard_death_SVM,
        hazard_death_SVI,
        hazard_death_ADI, 
        hazard_death_DCI) |> 
  ggplot()+
  geom_smooth(aes(x=SVM_PERCENTILE, y = hazard_death_SVM, color = "SVM"), data = avg_patient) +
    geom_smooth(aes(x=SVI_percentile, y = hazard_death_SVI, color = "SVI"), data = avg_patient1)+
    geom_smooth(aes(x=mean_ADI, y = hazard_death_ADI, color = "ADI"), data = avg_patient2)+
    geom_smooth(aes(x=dci_percentile, y = hazard_death_DCI, color = "DCI"), data = avg_patient3)+
  xlab("SDoH Composite Score Percentile")+
  ylab("Cause Specific HR of Death or Deterioration")+
  theme(legend.title = element_blank(),
        legend.position = "bottom")

spline_death
```


```{r}
predict_coxme <- function(object, 
                          newdata = NULL, 
                          type = c("lp", "risk"), 
                          se.fit = FALSE,
                          strata_ref = TRUE){
  
  if (!inherits(object, 'coxme'))
    stop("Primary argument much be a coxme object")
  
  type <- match.arg(type)
  n <- object$n[2]
  Terms <- delete.response(terms(object))
  has_strata <- !is.null(attr(Terms, "specials")$strata) 
  if (has_strata) 
    has_strata <- ifelse(length(attr(Terms, "specials")$strata) == 0, FALSE, has_strata)
  has_newdata  <- !is.null(newdata)
  
  if (!se.fit & type == "lp" & !has_newdata) return(object$linear.predictor)
  
  coef <- fixed.effects(object)
  mf <- survival:::model.frame.coxph(object)
  
  # boot.ci
  
  
  
  if (has_newdata){
    m <- model.frame(Terms, newdata)
  } else {
    m <- mf
  }
  
  # if strata update terms
  if (has_strata){
    strata_terms <- untangle.specials(Terms, "strata")
    Terms2 <- Terms[-strata_terms$terms]
  } else {
    Terms2 <- Terms
  }
  
  if (has_newdata){
    mm <- model.matrix(Terms2, m)
    mm <- mm[ ,-1]
  }
  
  # has strata and reference is strata
  # calculate strata means
  if (has_strata & strata_ref){
    # Full model matrix
    x <- model.matrix(Terms, data = mf)
    
    oldstrat <- mf[[strata_terms$vars]]
    xmeans <- rowsum(x, oldstrat)/as.vector(table(oldstrat))
  }
  
  if (!has_newdata){
    # extract all cols in x which matches Terms
    mm <- model.matrix(Terms2, data =mf)[ ,-1]
    m <- mf
  }
  
  if (has_strata & strata_ref){
    newstrat <- m[[strata_terms$vars]]
    mm <- mm - xmeans[match(newstrat, row.names(xmeans)), colnames(mm)]
  } else {
    mm <- mm - rep(object$means, each = nrow(m))
  }
  
  # if (!has_newdata & !has_strata){
  #   pred <- object$linear.predictor
  # }
  if (length(coef) == 1){
    pred <- mm * coef
  } else {
    pred <- (mm %*% coef)
  }
  
  
  if (se.fit) se <- sqrt(rowSums((mm %*% vcov(object)) * mm))
  if (type == "risk"){
    pred <- exp(pred)
    if (se.fit) se <- se * sqrt(pred)
  }
  if (se.fit) list(fit = pred, se.fit = se)
  else pred
}
```

