---
title: "Compare Deprivation Index Heart"
output: html_document
date: "2024-05-23"
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(dplyr)
library(survival)
library(haven)
library(lubridate)
library(ggsurvfit)
library(gtsummary)
library(readr)
library(consort)
library(splines)
library(readxl)
library(rms)
library(forcats)
library(cmprsk)
library(crrSC)
library(sf)
library(forestplot)
library(gridExtra)
library(pROC)
library(coxme)

```

## Loading Data

```{r data}
setwd("~/R/Parker")
cand_thor <- read_sas("cand_thor.sas7bdat", NULL)
slice_head(cand_thor,n=5)
```
## ASsigning Scores to Zip codes

## ADI

ADI is coded to the at the block level (12 digit FIPS), whereas the crosswalk document is only coded at the tract level (11 digit FIPS). Need to adjust accordingly. ADI from University of Wisconsin. Crosswalk document from US Department of Housing and Urban Dev.

```{r}
ADI <- read_csv("2021_ADI.csv")
ADI_crosswalk <-read_excel("11_digit_FIPS_to_ZIP_crosswalk.xlsx")

ADI$FIPS_11_digit <- substr(ADI$FIPS, 1, 11) 
ADI <- left_join(ADI,
                 select(ADI_crosswalk, TRACT, ZIP), 
                 by = c("FIPS_11_digit" = "TRACT") )
ADI <- ADI |> 
  group_by(ZIP) |> 
  mutate(mean_ADI = mean(as.numeric(ADI_NATRANK), na.rm = TRUE))
ADI |> 
  select(ZIP,  ADI_NATRANK, mean_ADI) |> 
  slice_head(n = 10)
ADI <- distinct(ADI, ZIP, .keep_all = TRUE)
rm(ADI_crosswalk)

```

## SVI

SVI is stupid and codes missing data as -999. 

```{r}
SVI <- read_csv("SVI_2022_US.csv")
SVI_crosswalk <- read_excel("11_digit_FIPS_to_ZIP_crosswalk.xlsx")

SVI <- left_join(SVI, select(SVI_crosswalk, TRACT, ZIP), by = c("FIPS" = "TRACT"))
SVI <- SVI |> 
  select(ZIP, RPL_THEMES) 
SVI <- SVI |> mutate(
  RPL_THEMES = case_when(RPL_THEMES < 0 ~ NA, .default = RPL_THEMES))
SVI <- SVI |> 
  group_by(ZIP) |> 
  mutate(SVI_percentile = mean(as.numeric(RPL_THEMES*100), na.rm = TRUE),
         median_SVI = median(as.numeric(RPL_THEMES*100), na.rm = TRUE))

SVI |>
  group_by(ZIP) |> 
  ggplot() +
  geom_point(mapping = aes(x = SVI_percentile, y = median_SVI))
SVI <- distinct(SVI, ZIP, .keep_all = TRUE)
rm(SVI_crosswalk)
slice_head(SVI, n=10) 


```

## SVM

```{r}
SVM_ZIP <- read_csv("SVM_ZIPCODE_SCORES_All States_All Counties_2024-01-23.csv")
SVM_ZIP$zipcode <- substr(SVM_ZIP$ZIPCODE, start = 1, stop = 5)
slice_head(SVM_ZIP,n= 5)

```
## DCI

```{r}
DCI <- read_excel("DCI-2017-2021-Scores-Only.xlsx")
DCI <- DCI |> 
  mutate(zip = as.character(`Zip Code`),
         dci_percentile = `2017-2021 Final Distress Score`)
slice_head(DCI, n = 5)
```


## Assign Deprivation Inidicies to Candidate ZIP Code

Good to know how many people have the 9 digit zip code

```{r}
cand_zip <- read_sas("canzip2403.sas7bdat", NULL)
cand_zip$ZIP = substr(cand_zip$CAN_PERM_ZIP,1,5)
cand_zip <- left_join(cand_zip, 
                              select(SVM_ZIP, ZIPCODE, SVM_SCORE, SVM_PERCENTILE),
                              by = c("ZIP" = "ZIPCODE"))
cand_zip <-left_join(cand_zip, SVI, by = c("ZIP" = "ZIP"))
cand_zip <- left_join(cand_zip, select(ADI, mean_ADI, ZIP), by = c("ZIP" = "ZIP"))
cand_zip <- left_join(cand_zip, select(DCI, dci_percentile, zip), by = c("ZIP" = "zip"))
rm(list = c("ADI", "SVI", "SVM_ZIP", "DCI"))

slice_head(cand_zip,n= 5)
```

## Subet Cohort

Only include patients added to the waitlist through 2019 or 2022.This allows us to observe everyone for at least 1 year this way. Only patients that are excluded are patients that are simultaneously listed for heart and lung.

```{r}
cand_thor <- cand_thor |> 
  subset(CAN_LISTING_DT > as.Date("2019-01-01") & CAN_LISTING_DT < as.Date("2022-12-31") & WL_ORG == "HR")
```

## VAD Type characterization

Credit to Kevin L. (https://github.com/kevinlazenby/post_transplant_survival_new_policy/blob/main/function_definitions.Rmd), lines 427 - 460

Durable LVADs (line 867) https://github.com/healthallocate/Heart-Continuous-Score/blob/main/6week_death_data_prep.Rmd


```{r}
cand_thor <- cand_thor |> 
  mutate(VAD = case_when(
        CAN_VAD_TY == 1 ~ paste0("None"),
        CAN_VAD_TY == 2 ~ paste0("LVAD"),
        CAN_VAD_TY == 3 ~ paste0("RVAD"),
        CAN_VAD_TY == 4 ~ paste0("TAH"),
        CAN_VAD_TY == 5 ~ paste0("LVAD+RVAD")),
      
      LVAD = case_when(
        VAD == "LVAD" ~ 1,
        TRUE ~ 0),
      
      durable_LVAD = case_when(
        CAN_VAD1 %in% c('202', '205', '206', '207', '208', '209', '210', 
                        '212', '213', '214', '223', '224', '233', '236', 
                        '239', '240', '312', '313', '314', '315', '316', '319', '322', 
                        '327', '330', '333', '334') | 
          CAN_VAD2 %in%  c('202', '205', '206', '207', '208', '209', '210', 
                        '212', '213', '214', '223', '224', '233', '236', 
                        '239', '240', '312', '313', '314', '315', '316', '319', '322', 
                        '327', '330', '333', '334') ~ 1,
        TRUE ~ 0),

      
      temporary_LVAD = case_when(
        durable_LVAD != 1 & !is.na(CAN_VAD1) ~ 1,
        TRUE ~ 0),
      
      RVAD = case_when(
        VAD == "RVAD" ~ 1,
        (CAN_VAD_TY != 1 & !is.na(CAN_VAD_TY) & 
       !(CAN_VAD1 %in% c('205', '236', '313', '330', '206', '208', '314', 
                         '210', '319', '216', '305', '217', '306', '223', 
                         '312', '224', '316', '230', '324', '231', '325', 
                         '232', '326', '233', '327'))) ~ 1,
        TRUE ~ 0),
      
      TAH = case_when(
        VAD == "TAH" ~ 1,
        TRUE ~ 0),
      
      LVAD_RVAD = case_when(
        VAD == "LVAD+RVAD" ~ 1,
        TRUE ~ 0))
  
      
```


## Intervals

Credit to Kevin Z. in USCRS score (lines 511 - 538)
https://github.com/healthallocate/Heart-Continuous-Score/blob/main/6week_death_data_prep.Rmd


```{r intervals}
 
cand_thor <- cand_thor |> 
  mutate(
    transplant_cand_thor = case_when(!is.na(REC_TX_DT) | CAN_REM_CD == '4' ~ 1,
      TRUE ~ 0),
    start_date = as.Date(CAN_LISTING_DT),
    transplant_date = as.Date(REC_TX_DT),
    removal_date = as.Date(CAN_REM_DT),
    last_active_date = as.Date(CAN_LAST_ACT_STAT_DT),
    last_inactive_date = as.Date(CAN_LAST_INACT_STAT_DT),
    death_date = as.Date(CAN_DEATH_DT),
    death_date_max = pmax(CAN_DEATH_DT, PERS_SSA_DEATH_DT, PERS_OPTN_DEATH_DT, na.rm=T)) %>%
  
  mutate(transplant_time = ifelse(!is.na(transplant_date), 
                                  as.numeric(transplant_date - start_date, units='days'), NA)) %>%
  mutate(death_time = ifelse(!is.na(death_date_max), as.numeric(death_date_max - start_date, units='days'), NA)) %>%
  mutate(removal_time = as.numeric(pmax(removal_date, last_active_date, last_inactive_date, na.rm = T) - 
                                     start_date, units='days'))
```

## Survival time

Survival time defined as time to transplant, death, or removal from waitlist, whichever comes first. Units = days (from previous code)

```{r survival time}
 cand_thor <- 
  cand_thor |> 
  mutate(
    survival_time = pmin(transplant_time, death_time, removal_time, na.rm = TRUE),
    survival_time = case_when(
      survival_time > 1095 ~ 1095,
      .default = survival_time
    ))
```

## Censoring

Death or deterioration on the waitlist coded as the event (1). Transplant is the censoring event (0).

```{r censoring}
cand_thor <- cand_thor |> 
  group_by(PX_ID) |> 
    mutate(status = case_when(
      survival_time == 1095 ~ 0, 
      survival_time == death_time ~ 2,
      CAN_REM_CD == 13 ~ 2,
      CAN_REM_CD %in% c(4, 14, 15, 18, 19, 20, 22, 23)  ~ 1,
      .default = 0
      ))

cand_thor$status <- factor(cand_thor$status,
                           levels = c(0,1,2),
                           labels = c("censor",  "transplant","death/deteriorate"))
```

Cause Specific Hazard setup

```{r}
cand_thor <- cand_thor |> 
  mutate(transplant_outcome = case_when(
    status == "censor" ~ 0,
    status == "transplant" ~ 1,
    .default = 0),
    death_outcome = case_when(
      status == "censor" ~ 0,
      status == "death/deteriorate" ~ 1,
      .default = 0
    ))
```

## STROBE/CONSORT Diagram

Inclusion and exclusion criteria Strobe diagram

```{r}
cand_thor <- cand_thor |> 
  mutate(eligibility = case_when(
    CAN_AGE_AT_LISTING < 18 ~ "Pediatric candidates",
    CAN_LISTING_DT == CAN_DEATH_DT ~ "Candidate listed and died same day",
  .default = NA),
    durable_LVAD_table1 = case_when(
        (is.na(eligibility) & durable_LVAD == 1) ~ "Durable LVAD",
        (is.na(eligibility) & durable_LVAD == 0) ~ "No Durable LVAD",
        .default = NA
      ),
  status_table1 = case_when(
    status =="censor" ~ "On waitlist at end of follow-up",
    status == "transplant" ~ "Received transplant",
    status == 'death/deteriorate' ~ "Died or removed from waitlist for deterioration"
  ))



txt <- gen_text(cand_thor$PX_ID, label = "Heart transplant candidate registrations, 2019-2022")
p_cons <- add_box(txt = txt)

txt <- gen_text(cand_thor$eligibility, label = "Excluded", bullet = TRUE)
p_cons <- add_side_box(p_cons, txt = txt)

p_cons <- add_box(p_cons, txt = gen_text(cand_thor$durable_LVAD_table1, label = "Adult candidate registrations"))

## Creating Analytic Sample 
cand_thor <- cand_thor |> 
  subset(is.na(eligibility))

txt <- gen_text(cand_thor$status_table1, bullet = TRUE)
p_cons <- add_box(p_cons, txt = txt, just = "left")


plot(p_cons, grViz = TRUE)
```


## Assigning Deprivation Indices based on Patient's permanent zip

Change to Person area code eventually

```{r center scores}
heart_SVM <- left_join(cand_thor,
                    select(cand_zip,SVM_SCORE, SVM_PERCENTILE, PX_ID, SVI_percentile, mean_ADI, dci_percentile), 
                    by = c("PX_ID" = "PX_ID"))
```

## Splitting SVM into Decile

```{r}
heart_SVM <- heart_SVM |> 
  group_by(PX_ID) |> 
  mutate(SVM_decile = 
           cut(SVM_PERCENTILE, c(-Inf, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100), labels = c("1st Decile", "2nd Decile", "3rd Decile", "4th Decile", "5th Decile", "6th Decile", "7th Decile", "8th Decile", "9th Decile", "10th Decile")),
           SVI_decile = cut(SVI_percentile, c(-Inf, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100), labels = c("1st Decile", "2nd Decile", "3rd Decile", "4th Decile", "5th Decile", "6th Decile", "7th Decile", "8th Decile", "9th Decile", "10th Decile")),
           ADI_decile = cut(mean_ADI,c(-Inf, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100), labels = c("1st Decile", "2nd Decile", "3rd Decile", "4th Decile", "5th Decile", "6th Decile", "7th Decile", "8th Decile", "9th Decile", "10th Decile")),
         DCI_decile = cut(dci_percentile,c(-Inf, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100), labels = c("1st Decile", "2nd Decile", "3rd Decile", "4th Decile", "5th Decile", "6th Decile", "7th Decile", "8th Decile", "9th Decile", "10th Decile")))
heart_SVM <- heart_SVM |> 
  mutate(svm_decile = as.numeric(SVM_decile),
        svi_decile = as.numeric(SVI_decile),
        adi_decile = as.numeric(ADI_decile),
        dci_decile = as.numeric(DCI_decile))
        
```



## Covariates

```{r covariates}
heart_SVM <- heart_SVM |> 
  mutate(can_race = case_when(
    CAN_RACE == "8" ~ "White",
    CAN_RACE == "16" ~ "Black or African American",
    CAN_RACE == "32" ~ "American Indian or Alaska Native",
    CAN_RACE == "64" ~ "Asian",
    CAN_RACE == "128" ~ "Native Hawaiian or Other Pacific Islander",
    CAN_RACE == "256" ~ "Arab or Middle Eastern",
    .default = "Other"),
    can_race = as.factor(can_race),
    ethnicity = case_when(
      CAN_ETHNICITY_SRTR == "NLATIN" ~ "Not Latinx",
      CAN_ETHNICITY_SRTR == "LATINO" ~ "Latinx"
      
    ),
    ethnicity = as.factor(ethnicity))
heart_SVM$can_race <- relevel(heart_SVM$can_race, ref = "White")
heart_SVM$ethnicity <- relevel(heart_SVM$ethnicity, ref = "Not Latinx")
heart_SVM <- heart_SVM |> 
  mutate(initial_status = case_when(
           CAN_INIT_STAT %in% c('1010', '1020', '1090', '1110', '2010', '2020', '2090', '2110', '3010', '6010', '6011', '6012', '9010') ~ "Status 1",
           CAN_INIT_STAT %in% c('1030', '1120', '2030', '2120', '6002', '6020', '6030', '9020') ~ "Status 2",
           CAN_INIT_STAT %in% c('1130', '2130', '6040') ~ "Status 3",
           CAN_INIT_STAT %in% c('1140', '2140', '6004') ~ "Status 4",
           CAN_INIT_STAT %in% c('1150', '2150') ~ "Status 5",
           CAN_INIT_STAT %in% c('1160', '2160') ~ "Status 6",
           .default = NA
         ))
```



## Table 1

```{r table 1}
heart_SVM |> 
  select(CAN_AGE_AT_LISTING, CAN_GENDER, can_race, durable_LVAD, temporary_LVAD, RVAD, LVAD_RVAD, TAH, initial_status, SVM_PERCENTILE, SVI_percentile, mean_ADI, dci_percentile, status_table1, CAN_ETHNICITY_SRTR, adi_decile ) |> 
  filter(adi_decile == 1 | adi_decile == 10) |> 
  group_by(PX_ID) |> 
  tbl_summary(by = adi_decile, label = list(
    can_race ~ "Race",
    CAN_GENDER ~ "Sex"
  ),
  statistic = list(all_continuous() ~ "{mean} ({sd})", all_categorical() ~ "{n} ({p}%)")) |> 
  modify_table_body(filter, !(variable == "PX_ID" ))

heart_SVM |> 
  select(CAN_AGE_AT_LISTING, CAN_GENDER, can_race, durable_LVAD, temporary_LVAD, RVAD, LVAD_RVAD, TAH, initial_status, SVM_PERCENTILE, SVI_percentile, mean_ADI, dci_percentile, status, CAN_ETHNICITY_SRTR ) |> 
  group_by(PX_ID) |> 
  tbl_summary(label = list(
    can_race ~ "Race",
    CAN_GENDER ~ "Sex"
  ),
  statistic = list(all_continuous() ~ "{mean} ({sd})", all_categorical() ~ "{n} ({p}%)"))
```

## Regression Models

SVM Transplant

```{r SVM crr transplant}
heart_SVM <- heart_SVM |> 
  mutate(sex = case_when(
    CAN_GENDER == "M" ~ 1,
    CAN_GENDER == "F" ~ 0),
    hispanic = case_when(
      CAN_ETHNICITY_SRTR == "LATINO" ~ 1,
      CAN_ETHNICITY_SRTR == "NLATIN" ~ 0,
      .default = NA),
    initial_status_numeric = case_when(
      CAN_INIT_STAT %in% c('1010', '1020', '1090', '1110', '2010', '2020', '2090', '2110', '3010', '6010', '6011', '6012', '9010') ~ 1,
           CAN_INIT_STAT %in% c('1030', '1120', '2030', '2120', '6002', '6020', '6030', '9020') ~ 2,
           CAN_INIT_STAT %in% c('1130', '2130', '6040') ~ 3,
           CAN_INIT_STAT %in% c('1140', '2140', '6004') ~ 4,
           CAN_INIT_STAT %in% c('1150', '2150') ~ 5,
           CAN_INIT_STAT %in% c('1160', '2160') ~ 6,
           .default = NA),
    status_numeric = case_when(
      status == "censor" ~ 0,
      status == "transplant" ~ 1,
      status == "death/deteriorate" ~ 2
    ))
```

## Spline Plots

Fail code for cumulative indidence is transplant. 2 cohorts: durable LVAD or no durable lvad

For prediction, set equal to the median age

Set ADI to be the case

objective of this is to equalize transplant across transplant. The absolute number of place based points, type of rating scale, shape of the rating scale will need to be determined through simulation optimization

Summer Gentry paper

```{r}
predict_coxme <- function(object, 
                          newdata = NULL, 
                          type = c("lp", "risk"), 
                          se.fit = FALSE,
                          strata_ref = TRUE){
  
  if (!inherits(object, 'coxme'))
    stop("Primary argument much be a coxme object")
  
  type <- match.arg(type)
  n <- object$n[2]
  Terms <- delete.response(terms(object))
  has_strata <- !is.null(attr(Terms, "specials")$strata) 
  if (has_strata) 
    has_strata <- ifelse(length(attr(Terms, "specials")$strata) == 0, FALSE, has_strata)
  has_newdata  <- !is.null(newdata)
  
  if (!se.fit & type == "lp" & !has_newdata) return(object$linear.predictor)
  
  coef <- fixed.effects(object)
  mf <- survival:::model.frame.coxph(object)
  
  # boot.ci
  
  
  
  if (has_newdata){
    m <- model.frame(Terms, newdata)
  } else {
    m <- mf
  }
  
  # if strata update terms
  if (has_strata){
    strata_terms <- untangle.specials(Terms, "strata")
    Terms2 <- Terms[-strata_terms$terms]
  } else {
    Terms2 <- Terms
  }
  
  if (has_newdata){
    mm <- model.matrix(Terms2, m)
    mm <- mm[ ,-1]
  }
  
  # has strata and reference is strata
  # calculate strata means
  if (has_strata & strata_ref){
    # Full model matrix
    x <- model.matrix(Terms, data = mf)
    
    oldstrat <- mf[[strata_terms$vars]]
    xmeans <- rowsum(x, oldstrat)/as.vector(table(oldstrat))
  }
  
  if (!has_newdata){
    # extract all cols in x which matches Terms
    mm <- model.matrix(Terms2, data =mf)[ ,-1]
    m <- mf
  }
  
  if (has_strata & strata_ref){
    newstrat <- m[[strata_terms$vars]]
    mm <- mm - xmeans[match(newstrat, row.names(xmeans)), colnames(mm)]
  } else {
    mm <- mm - rep(object$means, each = nrow(m))
  }
  
  # if (!has_newdata & !has_strata){
  #   pred <- object$linear.predictor
  # }
  if (length(coef) == 1){
    pred <- mm * coef
  } else {
    pred <- (mm %*% coef)
  }
  
  
  if (se.fit) se <- sqrt(rowSums((mm %*% vcov(object)) * mm))
  if (type == "risk"){
    pred <- exp(pred)
    if (se.fit) se <- se * sqrt(pred)
  }
  if (se.fit) list(fit = pred, se.fit = se)
  else pred
}
```



```{r}
set.seed(1234)
SVM_model_transplant <- coxme(Surv(survival_time,transplant_outcome) ~ rcs(SVM_PERCENTILE) + can_race + ethnicity +  CAN_GENDER + CAN_AGE_AT_LISTING + durable_LVAD +temporary_LVAD + RVAD + TAH + LVAD_RVAD + initial_status+ (1|CAN_LISTING_CTR_CD), data = heart_SVM)
SVI_model_transplant<- coxme(Surv(survival_time,transplant_outcome) ~ rcs(SVI_percentile) + can_race + ethnicity +  CAN_GENDER + CAN_AGE_AT_LISTING + durable_LVAD +temporary_LVAD + RVAD + TAH + LVAD_RVAD + initial_status +  (1|CAN_LISTING_CTR_CD), data = heart_SVM)
ADI_model_transplant <- coxme(Surv(survival_time,transplant_outcome) ~ rcs(mean_ADI) + can_race + ethnicity +  CAN_GENDER + CAN_AGE_AT_LISTING + durable_LVAD +temporary_LVAD + RVAD + TAH + LVAD_RVAD + initial_status + (1|CAN_LISTING_CTR_CD), data = heart_SVM)
DCI_model_transplant <- coxme(Surv(survival_time,transplant_outcome) ~ rcs(dci_percentile) + can_race + ethnicity +  CAN_GENDER + CAN_AGE_AT_LISTING + durable_LVAD +temporary_LVAD + RVAD + TAH + LVAD_RVAD + initial_status + (1|CAN_LISTING_CTR_CD), data = heart_SVM)


avg_patient <- expand_grid(
  SVM_PERCENTILE = seq(0,100),
  durable_LVAD = seq(0,1),
  can_race = c("White", "Black or African American", "Other", "Asian", "American Indian or Alaska Native", "Native Hawaiian or Other Pacific Islander"),
  initial_status = c("Status 1", "Status 2", "Status 3", "Status 4", "Status 5", "Status 6"),
  ethnicity = fct_unique(heart_SVM$ethnicity),
  CAN_GENDER = fct_unique(heart_SVM$CAN_GENDER),
  CAN_AGE_AT_LISTING = median(heart_SVM$CAN_AGE_AT_LISTING),
  temporary_LVAD = seq(0,1),
  RVAD = seq(0,1),
  TAH = seq(0,1),
  LVAD_RVAD = seq(0,1))
avg_patient1 <- expand_grid(
  SVI_percentile = seq(0,100),
  durable_LVAD = seq(0,1),
  can_race = c("White", "Black or African American", "Other", "Asian", "American Indian or Alaska Native", "Native Hawaiian or Other Pacific Islander"),
  initial_status = c("Status 1", "Status 2", "Status 3", "Status 4", "Status 5", "Status 6"),
  ethnicity = fct_unique(heart_SVM$ethnicity),
  CAN_GENDER = fct_unique(heart_SVM$CAN_GENDER),
  CAN_AGE_AT_LISTING = median(heart_SVM$CAN_AGE_AT_LISTING),
  temporary_LVAD = seq(0,1),
  RVAD = seq(0,1),
  TAH = seq(0,1),
  LVAD_RVAD = seq(0,1))
avg_patient2 <- expand_grid(
  mean_ADI = seq(0,100),
  durable_LVAD = seq(0,1),
  can_race = c("White", "Black or African American", "Other", "Asian", "American Indian or Alaska Native", "Native Hawaiian or Other Pacific Islander"),
  initial_status = c("Status 1", "Status 2", "Status 3", "Status 4", "Status 5", "Status 6"),
  ethnicity = fct_unique(heart_SVM$ethnicity),
  CAN_GENDER = fct_unique(heart_SVM$CAN_GENDER),
  CAN_AGE_AT_LISTING = median(heart_SVM$CAN_AGE_AT_LISTING),
  temporary_LVAD = seq(0,1),
  RVAD = seq(0,1),
  TAH = seq(0,1),
  LVAD_RVAD = seq(0,1))
avg_patient3 <- expand_grid(
  dci_percentile = seq(0,100),
  durable_LVAD = seq(0,1),
  can_race = c("White", "Black or African American", "Other", "Asian", "American Indian or Alaska Native", "Native Hawaiian or Other Pacific Islander"),
  initial_status = c("Status 1", "Status 2", "Status 3", "Status 4", "Status 5", "Status 6"),
  ethnicity = fct_unique(heart_SVM$ethnicity),
  CAN_GENDER = fct_unique(heart_SVM$CAN_GENDER),
  CAN_AGE_AT_LISTING = median(heart_SVM$CAN_AGE_AT_LISTING),
  temporary_LVAD = seq(0,1),
  RVAD = seq(0,1),
  TAH = seq(0,1),
  LVAD_RVAD = seq(0,1))



hazard_transplant_SVM <- predict_coxme(SVM_model_transplant, newdata = avg_patient, type=c("risk"))
hazard_transplant_SVI <- predict_coxme(SVI_model_transplant, newdata = avg_patient1, type = c("risk"))
hazard_transplant_ADI <- predict_coxme(ADI_model_transplant, newdata = avg_patient2, type = c("risk"))
hazard_transplant_DCI <- predict_coxme(DCI_model_transplant, newdata = avg_patient3, type = c("risk"))


spline_transplant <- avg_patient |> 
  cbind(hazard_transplant_SVM,
        hazard_transplant_SVI,
        hazard_transplant_ADI, 
        hazard_transplant_DCI) |> 
  ggplot()+
  geom_smooth(aes(x=SVM_PERCENTILE, y = hazard_transplant_SVM, color = "SVM"), data = avg_patient) +
    geom_smooth(aes(x=SVI_percentile, y = hazard_transplant_SVI, color = "SVI"), data = avg_patient1)+
    geom_smooth(aes(x=mean_ADI, y = hazard_transplant_ADI, color = "ADI"), data = avg_patient2)+
    geom_smooth(aes(x=dci_percentile, y = hazard_transplant_DCI, color = "DCI"), data = avg_patient3)+
  xlab("SDoH Composite Score Percentile")+
  ylab("Cause Specific HR of Transplant")+
  theme(legend.title = element_blank(),
        legend.position = "bottom")

spline_transplant

```

```{r}
set.seed(1234)
SVM_model_death<- coxme(Surv(survival_time,death_outcome) ~ rcs(SVM_PERCENTILE) + can_race + ethnicity +  CAN_GENDER + CAN_AGE_AT_LISTING + durable_LVAD +temporary_LVAD + RVAD + TAH + LVAD_RVAD + initial_status+ (1|CAN_LISTING_CTR_CD), data = heart_SVM)
SVI_model_death<- coxme(Surv(survival_time,death_outcome) ~ rcs(SVI_percentile) + can_race + ethnicity +  CAN_GENDER + CAN_AGE_AT_LISTING + durable_LVAD +temporary_LVAD + RVAD + TAH + LVAD_RVAD + initial_status +  (1|CAN_LISTING_CTR_CD), data = heart_SVM)
ADI_model_death <- coxme(Surv(survival_time,death_outcome) ~ rcs(mean_ADI) + can_race + ethnicity +  CAN_GENDER + CAN_AGE_AT_LISTING + durable_LVAD +temporary_LVAD + RVAD + TAH + LVAD_RVAD + initial_status + (1|CAN_LISTING_CTR_CD), data = heart_SVM)
DCI_model_death <- coxme(Surv(survival_time,death_outcome) ~ rcs(dci_percentile) + can_race + ethnicity +  CAN_GENDER + CAN_AGE_AT_LISTING + durable_LVAD +temporary_LVAD + RVAD + TAH + LVAD_RVAD + initial_status + (1|CAN_LISTING_CTR_CD), data = heart_SVM)


avg_patient <- expand_grid(
  SVM_PERCENTILE = seq(0,100),
  durable_LVAD = seq(0,1),
  can_race = c("White", "Black or African American", "Other", "Asian", "American Indian or Alaska Native", "Native Hawaiian or Other Pacific Islander"),
  initial_status = c("Status 1", "Status 2", "Status 3", "Status 4", "Status 5", "Status 6"),
  ethnicity = fct_unique(heart_SVM$ethnicity),
  CAN_GENDER = fct_unique(heart_SVM$CAN_GENDER),
  CAN_AGE_AT_LISTING = median(heart_SVM$CAN_AGE_AT_LISTING),
  temporary_LVAD = seq(0,1),
  RVAD = seq(0,1),
  TAH = seq(0,1),
  LVAD_RVAD = seq(0,1))
avg_patient1 <- expand_grid(
  SVI_percentile = seq(0,100),
  durable_LVAD = seq(0,1),
  can_race = c("White", "Black or African American", "Other", "Asian", "American Indian or Alaska Native", "Native Hawaiian or Other Pacific Islander"),
  initial_status = c("Status 1", "Status 2", "Status 3", "Status 4", "Status 5", "Status 6"),
  ethnicity = fct_unique(heart_SVM$ethnicity),
  CAN_GENDER = fct_unique(heart_SVM$CAN_GENDER),
  CAN_AGE_AT_LISTING = median(heart_SVM$CAN_AGE_AT_LISTING),
  temporary_LVAD = seq(0,1),
  RVAD = seq(0,1),
  TAH = seq(0,1),
  LVAD_RVAD = seq(0,1))
avg_patient2 <- expand_grid(
  mean_ADI = seq(0,100),
  durable_LVAD = seq(0,1),
  can_race = c("White", "Black or African American", "Other", "Asian", "American Indian or Alaska Native", "Native Hawaiian or Other Pacific Islander"),
  initial_status = c("Status 1", "Status 2", "Status 3", "Status 4", "Status 5", "Status 6"),
  ethnicity = fct_unique(heart_SVM$ethnicity),
  CAN_GENDER = fct_unique(heart_SVM$CAN_GENDER),
  CAN_AGE_AT_LISTING = median(heart_SVM$CAN_AGE_AT_LISTING),
  temporary_LVAD = seq(0,1),
  RVAD = seq(0,1),
  TAH = seq(0,1),
  LVAD_RVAD = seq(0,1))
avg_patient3 <- expand_grid(
  dci_percentile = seq(0,100),
  durable_LVAD = seq(0,1),
  can_race = c("White", "Black or African American", "Other", "Asian", "American Indian or Alaska Native", "Native Hawaiian or Other Pacific Islander"),
  initial_status = c("Status 1", "Status 2", "Status 3", "Status 4", "Status 5", "Status 6"),
  ethnicity = fct_unique(heart_SVM$ethnicity),
  CAN_GENDER = fct_unique(heart_SVM$CAN_GENDER),
  CAN_AGE_AT_LISTING = median(heart_SVM$CAN_AGE_AT_LISTING),
  temporary_LVAD = seq(0,1),
  RVAD = seq(0,1),
  TAH = seq(0,1),
  LVAD_RVAD = seq(0,1))


hazard_death_SVM <- predict_coxme(SVM_model_death, newdata = avg_patient, type = c("risk"))
hazard_death_SVI <- predict_coxme(SVI_model_death, newdata = avg_patient1, type = c("risk"))
hazard_death_ADI <- predict_coxme(ADI_model_death, newdata = avg_patient2, type = c("risk"))
hazard_death_DCI <- predict_coxme(DCI_model_death, newdata = avg_patient3, type = c("risk"))



spline_death <- avg_patient |> 
  cbind(hazard_death_SVM,
        hazard_death_SVI,
        hazard_death_ADI, 
        hazard_death_DCI) |> 
  ggplot()+
  geom_smooth(aes(x=SVM_PERCENTILE, y = hazard_death_SVM, color = "SVM"), data = avg_patient) +
    geom_smooth(aes(x=SVI_percentile, y = hazard_death_SVI, color = "SVI"), data = avg_patient1)+
    geom_smooth(aes(x=mean_ADI, y = hazard_death_ADI, color = "ADI"), data = avg_patient2)+
    geom_smooth(aes(x=dci_percentile, y = hazard_death_DCI, color = "DCI"), data = avg_patient3)+
  xlab("SDoH Composite Score Percentile")+
  ylab("Cause Specific HR of Death or Deterioration")+
  theme(legend.title = element_blank(),
        legend.position = "bottom")

spline_death
```
